---
title: "differential expression between clusters"
format: 
  html:
    code-fold: true
editor: visual
toc: true
execute:
  warning: false
  message: false
---

# differential expression between clusters

In this notebook we will assess if there are differentially expressed genes between d8 and d60 between individual conditions

As a test we will start of with comparing d8 MuHV44 vs d60 MuHV4 and d8 PR8 vs d60 PR8

## load in the data and the required packages

```{r}

easypackages::libraries("viridis","patchwork", "Seurat", "tidyverse","tidyseurat", "DESeq2", "ggrepel", "writexl", "clusterProfiler", "DOSE", "pathview", "biomaRt", "org.Mm.eg.db", "enrichplot")
obj.v5 <- read_rds("../../Documents/machiels_lab_viral/intermediate_data/seurat_obj_central.rds")

```

## data manipulations

```{r}
seurat_nod8 <- obj.v5 %>%filter(!str_detect(day_mock, "d8"))
seurat_nomock <- obj.v5 %>% filter(!str_detect(day_mock, "Mock"))
seurat_nomock = seurat_nomock %>% filter(!str_detect(condition, "Mock"))
#d60
seurat_d60 <- seurat_nod8 %>%filter(str_detect(day_mock, "d60"))
seurat_onlyMuHV4_d60 <- seurat_d60 %>% filter(str_detect(condition, "MuHV4"))
seurat_onlyMuHV4_d60_bal <- seurat_onlyMuHV4_d60 %>%  filter(str_detect(sample_type, "bal"))

#mock
seurat_mock <- seurat_nod8 %>%filter(str_detect(day_mock, "Mock"))
seurat_mock_onlyMuHV4 = seurat_mock %>%  filter(str_detect(condition, "MuHV4"))
seurat_mock_onlyMuHV4_bal = seurat_mock_onlyMuHV4 %>% filter(str_detect(sample_type, "bal"))


seurat_nomock_onlyMuHV4 = seurat_nomock %>% filter(str_detect(condition, "MuHV4"))
seurat_nomock_onlyMuHV4_bal = seurat_nomock_onlyMuHV4 %>% filter(str_detect(sample_type, "bal"))
seurat_nomock_onlyMuHV4_lung = seurat_nomock_onlyMuHV4 %>% filter(str_detect(sample_type, "lung"))

seurat_nomock_onlyPR8 = seurat_nomock %>% filter(str_detect(condition, "PR8"))
seurat_nomock_onlyPR8_bal = seurat_nomock_onlyPR8 %>% filter(str_detect(sample_type, "bal"))
seurat_nomock_onlyPR8_lung = seurat_nomock_onlyPR8 %>% filter(str_detect(sample_type, "lung"))
DefaultAssay(seurat_nomock_onlyMuHV4 )= "RNA"
DefaultAssay(seurat_nomock_onlyPR8 )= "RNA"

```

## 1) Exploratory analysis d8 vs d60 MuHV4

### CD11c

```{r}
featureplot_d8d60_MuHV4_cd11c = FeaturePlot(seurat_nomock_onlyMuHV4, reduction = "umap.harmony_8dims", label = T, split.by = c("day_mock"), features = "Itgax", repel = TRUE)


featureplot_d8d60_MuHV4_cd11c 

# + ggtitle("",subtitle = "UMAP plots with expression of CD11c at day 60 between Ms4a3 conditions in  mouse bal") + theme(plot.title = element_text(hjust = 0.5))



ggsave(filename = "featureplot_d8d60_MuHV4_cd11c.png", width = 45, height = 20, units = "cm", path = "../../Documents/machiels_lab_viral/output/plotsUmap/explorationDEA/")
```

### PPARg

```{r}
featureplot_d8d60_MuHV4_pparg = FeaturePlot(seurat_nomock_onlyMuHV4, reduction = "umap.harmony_8dims", label = T, split.by = c("day_mock"), features = "Pparg", repel = TRUE)

featureplot_d8d60_MuHV4_pparg 

ggsave(filename = "featureplot_d8d60_MuHV4_pparg.png", width = 45, height = 20, units = "cm", path =  "../../Documents/machiels_lab_viral/output/plotsUmap/explorationDEA/")
```

### SiglecF

```{r}
featureplot_d8d60_MuHV4_siglecf = FeaturePlot(seurat_nomock_onlyMuHV4, reduction = "umap.harmony_8dims", label = T, split.by = c("day_mock"), features = "Siglecf", repel = TRUE)


featureplot_d8d60_MuHV4_siglecf


ggsave(filename = "featureplot_d8d60_MuHV4_siglecf.png", width = 45, height = 20, units = "cm", path =  "../../Documents/machiels_lab_viral/output/plotsUmap/explorationDEA/")

```

### H2-ia-ie

```{r}
featureplot_d8d60_MuHV4_h2 = FeaturePlot(seurat_nomock_onlyMuHV4, reduction = "umap.harmony_8dims", label = T, split.by = c("day_mock"), features = "H2-ia-ie-AbSeq", repel = TRUE)


featureplot_d8d60_MuHV4_h2

# + ggtitle("",subtitle = "UMAP plots with expression of CD11c at day 60 between Ms4a3 conditions in  mouse bal") + theme(plot.title = element_text(hjust = 0.5))



ggsave(filename = "featureplot_d8d60_MuHV4_h2.png", width = 45, height = 20, units = "cm", path =  "../../Documents/machiels_lab_viral/output/plotsUmap/explorationDEA/")
```

### CD274

```{r}
featureplot_d8d60_MuHV4_CD274 = FeaturePlot(seurat_nomock_onlyMuHV4, reduction = "umap.harmony_8dims", label = T, split.by = c("day_mock"), features = "Cd274", repel = TRUE)


featureplot_d8d60_MuHV4_CD274



ggsave(filename = "featureplot_d8d60_MuHV4_cd274.png", width = 45, height = 20, units = "cm", path =  "../../Documents/machiels_lab_viral/output/plotsUmap/explorationDEA/")
```

## 1 A) Differential expression analysis MuHV4 d8 vs d60

```{r}
Idents(seurat_nomock_onlyMuHV4) <- "day_mock"
#  every gene in your dataset has a zero count in at least one sample.

markers_MuHV4_d60 <- FindMarkers(seurat_nomock_onlyMuHV4 ,
                                 ident.1 =  "d60",
                                 ident.2 =  "d8",
                                 test.use = "wilcox_limma", 
                                 min.pct = 0.2) |> as_tibble(rownames="gene")
markers_MuHV4_d8 <- FindMarkers(seurat_nomock_onlyMuHV4 ,
                                 ident.1 =  "d8",
                                 ident.2 =  "d60",
                                 test.use = "wilcox_limma", 
                                 min.pct = 0.2) |> as_tibble(rownames="gene")
                                 

```

### export the raw and filtered

```{r}
markers_MuHV4_d60_filtered = markers_MuHV4_d60[markers_MuHV4_d60$avg_log2FC > 1,]
nonsignficant_p_adj_d60MuHV4 <- which(markers_MuHV4_d60_filtered$p_val_adj <= 0.05)
markers_MuHV4_d60_filtered <- markers_MuHV4_d60_filtered[nonsignficant_p_adj_d60MuHV4, ]

write_xlsx(markers_MuHV4_d60, "../../Documents/machiels_lab_viral/output/DEA/markers_MuHV4_d60.xlsx")
write_xlsx(markers_MuHV4_d60_filtered, "../../Documents/machiels_lab_viral/output/DEA/markers_MuHV4_d60_filtered.xlsx")

markers_MuHV4_d8_filtered = markers_MuHV4_d8[markers_MuHV4_d8$avg_log2FC > 1,]
nonsignficant_p_adj_d8MuHV4 <- which(markers_MuHV4_d8_filtered$p_val_adj <= 0.05)
markers_MuHV4_d8_filtered <- markers_MuHV4_d8_filtered[nonsignficant_p_adj_d8MuHV4, ]

write_xlsx(markers_MuHV4_d8, "../../Documents/machiels_lab_viral/output/DEA/markers_MuHV4_d8.xlsx")
write_xlsx(markers_MuHV4_d8_filtered, "../../Documents/machiels_lab_viral/output/DEA/markers_MuHV4_d8_filtered.xlsx")

```

### volcano plot Muhv4 d60

```{r}
markers_MuHV4_d60$diffexpressed <- "NO"
markers_MuHV4_d60$diffexpressed[markers_MuHV4_d60$avg_log2FC > 1] <- "UP"

markers_MuHV4_d60$diffexpressed[markers_MuHV4_d60$avg_log2FC < -1] <- "DOWN"

markers_MuHV4_d60$delabel <- NA

markers_MuHV4_d60$delabel[markers_MuHV4_d60$diffexpressed != "NO"]<-markers_MuHV4_d60$gene[markers_MuHV4_d60$diffexpressed != "NO"]



volcano_muhv4_d60 <- ggplot(data=markers_MuHV4_d60,
                           aes(x=avg_log2FC, y=-log10(p_val_adj), col = diffexpressed,                            label = delabel)) +
                           geom_point() + theme_minimal() +
                           scale_x_continuous(breaks = seq(floor(min (markers_MuHV4_d60$avg_log2FC)), 
                           ceiling(max(markers_MuHV4_d60$avg_log2FC)), by = 0.5)) +
                           geom_text_repel(max.overlaps = 100)
volcano_muhv4_d60
ggsave("volcano_MuHV4_d60.png",width = 45, height = 20, units = "cm", path = "../../Documents/machiels_lab_viral/output/DEA/volcano/")

```

### volcano plot Muhv4 d8

```{r}
markers_MuHV4_d8$diffexpressed <- "NO"
markers_MuHV4_d8$diffexpressed[markers_MuHV4_d8$avg_log2FC > 1] <- "UP"

markers_MuHV4_d8$diffexpressed[markers_MuHV4_d8$avg_log2FC < -1] <- "DOWN"

markers_MuHV4_d8$delabel <- NA

markers_MuHV4_d8$delabel[markers_MuHV4_d8$diffexpressed != "NO"]<-markers_MuHV4_d8$gene[markers_MuHV4_d8$diffexpressed != "NO"]



volcano_muhv4_d8 <- ggplot(data=markers_MuHV4_d8,
                           aes(x=avg_log2FC, y=-log10(p_val_adj), col = diffexpressed,                            label = delabel)) +
                           geom_point() + theme_minimal() +
                           scale_x_continuous(breaks = seq(floor(min (markers_MuHV4_d8$avg_log2FC)), 
                           ceiling(max(markers_MuHV4_d8$avg_log2FC)), by = 0.5)) +
                           geom_text_repel(max.overlaps = 100)
volcano_muhv4_d8
ggsave("volcano_MuHV4_d8.png",width = 45, height = 20, units = "cm", path = "../../Documents/machiels_lab_viral/output/DEA/volcano/")
```

## 1 B) Functional enrichment + pathway analysis of marker genes of Muhv4 d60

```{r}
genes_MuHV4_d60 = markers_MuHV4_d60_filtered$gene
entrezids_MuHV4_d60_markers = select(org.Mm.eg.db, keys=genes_MuHV4_d60, keytype="SYMBOL", columns="ENTREZID")
entrezids_MuHV4_d60_markers = na.omit(entrezids_MuHV4_d60_markers)
```

```{r}
all_MuHV4_d60_genes = as.character(markers_MuHV4_d60$gene)
entrezids_MuHV4_d60_all = select(org.Mm.eg.db, keys=all_MuHV4_d60_genes, keytype="SYMBOL", columns="ENTREZID")
entrezids_MuHV4_d60_all = na.omit(entrezids_MuHV4_d60_all)
```

```{r}
GO_enrichment_MuHV4_d60 = enrichGO(gene = entrezids_MuHV4_d60_markers$ENTREZID,
                                   universe = entrezids_MuHV4_d60_all$ENTREZID,
                                   OrgDb = org.Mm.eg.db, 
                                   ont = "BP", 
                                   qvalueCutoff = 0.05, 
                                   readable = TRUE)
```

```{r}
cluster_summary_enrichment_MuHV4_d60 = data.frame(GO_enrichment_MuHV4_d60)
write.csv(cluster_summary_enrichment_MuHV4_d60, "../../Documents/machiels_lab_viral/output/DEA/clusterProfiler/MuHV4_d60/results_clusterProfiler_MuHV4_d60.csv")

```

```{r}
dotplot_goEnrichment_MuHV4_d60 = dotplot(GO_enrichment_MuHV4_d60, showCategory=50)
ggsave(filename = "dotplot_goEnrichment_MuHV4_d60.png", width = 45, height = 50, units = "cm", path =  "../../Documents/machiels_lab_viral/output/DEA/clusterProfiler/MuHV4_d60/")
```

```{r}
emapplot_goEnrichment_MuHV4_d60 = pairwise_termsim(GO_enrichment_MuHV4_d60)
emapplot_goEnrichment_MuHV4_d60 = emapplot(emapplot_goEnrichment_MuHV4_d60)


ggsave(filename = "emapplot_goEnrichment_MuHV4_d60.png", width = 45, height = 50, units = "cm", path =  "../../Documents/machiels_lab_viral/output/DEA/clusterProfiler/MuHV4_d60/")
```

```{r}
muhv4_d60_logfold = markers_MuHV4_d60_filtered$avg_log2FC
names(muhv4_d60_logfold) = markers_MuHV4_d60_filtered$gene

cnetplot(GO_enrichment_MuHV4_d60,
         categorySize="pvalue", 
         showCategory = 5, 
         foldChange=muhv4_d60_logfold, 
         vertex.label.font=6)

ggsave(filename = "cnetplot_goEnrichment_MuHV4_d60.png", width = 45, height = 50, units = "cm", path =  "../../Documents/machiels_lab_viral/output/DEA/clusterProfiler/MuHV4_d60/")
```

## 1 C) Functional enrichment + pathway analysis of marker genes of Muhv4 d8

```{r}
genes_MuHV4_d8 = markers_MuHV4_d8_filtered$gene
entrezids_MuHV4_d8_markers = select(org.Mm.eg.db, keys=genes_MuHV4_d8, keytype="SYMBOL", columns="ENTREZID")
entrezids_MuHV4_d8_markers = na.omit(entrezids_MuHV4_d8_markers)
```

```{r}
all_MuHV4_d8_genes = as.character(markers_MuHV4_d8$gene)
entrezids_MuHV4_d8_all = select(org.Mm.eg.db, keys=all_MuHV4_d8_genes, keytype="SYMBOL", columns="ENTREZID")
entrezids_MuHV4_d8_all = na.omit(entrezids_MuHV4_d8_all)
```

```{r}
GO_enrichment_MuHV4_d8 = enrichGO(gene = entrezids_MuHV4_d8_markers$ENTREZID,
                                   universe = entrezids_MuHV4_d8_all$ENTREZID,
                                   OrgDb = org.Mm.eg.db, 
                                   ont = "BP", 
                                   qvalueCutoff = 0.05, 
                                   readable = TRUE)
```

```{r}
cluster_summary_enrichment_MuHV4_d8 = data.frame(GO_enrichment_MuHV4_d8)
write.csv(cluster_summary_enrichment_MuHV4_d8, "../../Documents/machiels_lab_viral/output/DEA/clusterProfiler/MuHV4_d8/results_clusterProfiler_MuHV4_d8.csv")
```

```{r}
dotplot_goEnrichment_MuHV4_d8 = dotplot(GO_enrichment_MuHV4_d8, showCategory=50)
ggsave(filename = "dotplot_goEnrichment_MuHV4_d8.png", width = 45, height = 50, units = "cm", path =  "../../Documents/machiels_lab_viral/output/DEA/clusterProfiler/MuHV4_d8/")
```

shows the relationship between the top 50 most significantly enriched GO terms (padj.), by grouping similar terms together. The color represents the p-values relative to the other displayed terms (brighter red is more significant) and the size of the terms represents the number of genes that are significant from our list.

```{r}
emapplot_goEnrichment_MuHV4_d8 = pairwise_termsim(GO_enrichment_MuHV4_d8)
emapplot_goEnrichment_MuHV4_d8 = emapplot(emapplot_goEnrichment_MuHV4_d8)


ggsave(filename = "emapplot_goEnrichment_MuHV4_d8.png", width = 45, height = 50, units = "cm", path =  "../../Documents/machiels_lab_viral/output/DEA/clusterProfiler/MuHV4_d8/")
```

```{r}
muhv4_d8_logfold = markers_MuHV4_d8_filtered$avg_log2FC
names(muhv4_d8_logfold) = markers_MuHV4_d8_filtered$gene

cnetplot(GO_enrichment_MuHV4_d8,
         categorySize="pvalue", 
         showCategory = 5, 
         foldChange=muhv4_d8_logfold, 
         vertex.label.font=6)

ggsave(filename = "cnetplot_goEnrichment_MuHV4_d8.png", width = 45, height = 50, units = "cm", path =  "../../Documents/machiels_lab_viral/output/DEA/clusterProfiler/MuHV4_d8/")
```

## 2) Exploratory analysis d8 vs d60 PR8

### CD11C

```{r}
featureplot_d8d60_PR8_cd11c = FeaturePlot(seurat_nomock_onlyPR8, reduction = "umap.harmony_8dims", label = T, split.by = c("day_mock"), features = "Itgax", repel = TRUE)


featureplot_d8d60_PR8_cd11c 

# + ggtitle("",subtitle = "UMAP plots with expression of CD11c at day 60 between Ms4a3 conditions in  mouse bal") + theme(plot.title = element_text(hjust = 0.5))



ggsave(filename = "featureplot_d8d60_PR8_cd11c.png", width = 45, height = 50, units = "cm", path = "../../Documents/machiels_lab_viral/output/plotsUmap/explorationDEA/")
```

### PRARg

```{r}
featureplot_d8d60_PR8_pparg = FeaturePlot(seurat_nomock_onlyPR8, reduction = "umap.harmony_8dims", label = T, split.by = c("day_mock"), features = "Pparg", repel = TRUE)

featureplot_d8d60_PR8_pparg 

ggsave(filename = "featureplot_d8d60_PR8_pparg.png", width = 45, height = 20, units = "cm", path =  "../../Documents/machiels_lab_viral/output/plotsUmap/explorationDEA/")
```

### SiglecF

```{r}
featureplot_d8d60_PR8_siglecf = FeaturePlot(seurat_nomock_onlyPR8, reduction = "umap.harmony_8dims", label = T, split.by = c("day_mock"), features = "Siglecf", repel = TRUE)


featureplot_d8d60_PR8_siglecf


ggsave(filename = "featureplot_d8d60_PR8_siglecf.png", width = 45, height = 20, units = "cm", path =  "../../Documents/machiels_lab_viral/output/plotsUmap/explorationDEA/")
```

### H2-ia-ie

```{r}
featureplot_d8d60_PR8_h2 = FeaturePlot(seurat_nomock_onlyPR8, reduction = "umap.harmony_8dims", label = T, split.by = c("day_mock"), features = "H2-ia-ie-AbSeq", repel = TRUE)


featureplot_d8d60_PR8_h2

# + ggtitle("",subtitle = "UMAP plots with expression of CD11c at day 60 between Ms4a3 conditions in  mouse bal") + theme(plot.title = element_text(hjust = 0.5))



ggsave(filename = "featureplot_d8d60_PR8_h2.png", width = 45, height = 20, units = "cm", path =  "../../Documents/machiels_lab_viral/output/plotsUmap/explorationDEA/")
```

### CD274

```{r}
featureplot_d8d60_PR8_CD274 = FeaturePlot(seurat_nomock_onlyPR8, reduction = "umap.harmony_8dims", label = T, split.by = c("day_mock"), features = "Cd274", repel = TRUE)


featureplot_d8d60_PR8_CD274



ggsave(filename = "featureplot_d8d60_PR8_cd274.png", width = 45, height = 20, units = "cm", path =  "../../Documents/machiels_lab_viral/output/plotsUmap/explorationDEA/")
```

## 2 A) Differential expression analysis PR8 d8 vs d60

```{r}
Idents(seurat_nomock_onlyPR8) <- "day_mock"
#  every gene in your dataset has a zero count in at least one sample.

markers_PR8_d60 <- FindMarkers(seurat_nomock_onlyPR8 ,
                                 ident.1 =  "d60",
                                 ident.2 =  "d8",
                                 test.use = "wilcox_limma", 
                                 min.pct = 0.2) |> as_tibble(rownames="gene")
markers_PR8_d8 <- FindMarkers(seurat_nomock_onlyPR8 ,
                                 ident.1 =  "d8",
                                 ident.2 =  "d60",
                                 test.use = "wilcox_limma", 
                                 min.pct = 0.2) |> as_tibble(rownames="gene")
```

### export the raw and filtered

```{r}
markers_PR8_d60_filtered = markers_PR8_d60[markers_PR8_d60$avg_log2FC > 1,]
nonsignficant_p_adj_d60PR8 <- which(markers_PR8_d60_filtered$p_val_adj <= 0.05)
markers_PR8_d60_filtered <- markers_PR8_d60_filtered[nonsignficant_p_adj_d60PR8, ]

write_xlsx(markers_PR8_d60, "../../Documents/machiels_lab_viral/output/DEA/markers_PR8_d60.xlsx")
write_xlsx(markers_PR8_d60_filtered, "../../Documents/machiels_lab_viral/output/DEA/markers_PR8_d60_filtered.xlsx")

markers_PR8_d8_filtered = markers_PR8_d8[markers_PR8_d8$avg_log2FC > 1,]
nonsignficant_p_adj_d8PR8 <- which(markers_PR8_d8_filtered$p_val_adj <= 0.05)
markers_PR8_d8_filtered <- markers_PR8_d8_filtered[nonsignficant_p_adj_d8PR8, ]

write_xlsx(markers_PR8_d8, "../../Documents/machiels_lab_viral/output/DEA/markers_PR8_d8.xlsx")
write_xlsx(markers_PR8_d8_filtered, "../../Documents/machiels_lab_viral/output/DEA/markers_PR8_d8_filtered.xlsx")
```

### volcano plot PR8 d60

```{r}
markers_PR8_d60$diffexpressed <- "NO"
markers_PR8_d60$diffexpressed[markers_PR8_d60$avg_log2FC > 1] <- "UP"

markers_PR8_d60$diffexpressed[markers_PR8_d60$avg_log2FC < -1] <- "DOWN"

markers_PR8_d60$delabel <- NA

markers_PR8_d60$delabel[markers_PR8_d60$diffexpressed != "NO"]<-markers_PR8_d60$gene[markers_PR8_d60$diffexpressed != "NO"]



volcano_PR8_d60 <- ggplot(data=markers_PR8_d60,
                           aes(x=avg_log2FC, y=-log10(p_val_adj), col = diffexpressed,                            label = delabel)) +
                           geom_point() + theme_minimal() +
                           scale_x_continuous(breaks = seq(floor(min (markers_PR8_d60$avg_log2FC)), 
                           ceiling(max(markers_PR8_d60$avg_log2FC)), by = 0.5)) +
                           geom_text_repel(max.overlaps = 100)
volcano_PR8_d60
ggsave("volcano_PR8_d60.png",width = 45, height = 20, units = "cm", path = "../../Documents/machiels_lab_viral/output/DEA/volcano/")
```

### volcano plot PR8 d8

```{r}
markers_PR8_d8$diffexpressed <- "NO"
markers_PR8_d8$diffexpressed[markers_PR8_d8$avg_log2FC > 1] <- "UP"

markers_PR8_d8$diffexpressed[markers_PR8_d8$avg_log2FC < -1] <- "DOWN"

markers_PR8_d8$delabel <- NA

markers_PR8_d8$delabel[markers_PR8_d8$diffexpressed != "NO"]<-markers_PR8_d8$gene[markers_PR8_d8$diffexpressed != "NO"]



volcano_PR8_d8 <- ggplot(data=markers_PR8_d8,
                           aes(x=avg_log2FC, y=-log10(p_val_adj), col = diffexpressed,                            label = delabel)) +
                           geom_point() + theme_minimal() +
                           scale_x_continuous(breaks = seq(floor(min (markers_PR8_d8$avg_log2FC)), 
                           ceiling(max(markers_PR8_d8$avg_log2FC)), by = 0.5)) +
                           geom_text_repel(max.overlaps = 100)
volcano_PR8_d8
ggsave("volcano_PR8_d8.png",width = 45, height = 20, units = "cm", path = "../../Documents/machiels_lab_viral/output/DEA/volcano/")
```

## 2 B) Functional enrichment + pathway analysis of marker genes of PR8 d60 

```{r}
genes_PR8_d60 = markers_PR8_d60_filtered$gene
entrezids_PR8_d60_markers = select(org.Mm.eg.db, keys=genes_PR8_d60, keytype="SYMBOL", columns="ENTREZID")
entrezids_PR8_d60_markers = na.omit(entrezids_PR8_d60_markers)
```

```{r}
all_PR8_d60_genes = as.character(markers_PR8_d60$gene)
entrezids_PR8_d60_all = select(org.Mm.eg.db, keys=all_PR8_d60_genes, keytype="SYMBOL", columns="ENTREZID")
entrezids_PR8_d60_all = na.omit(entrezids_PR8_d60_all)
```

```{r}
GO_enrichment_PR8_d60 = enrichGO(gene = entrezids_PR8_d60_markers$ENTREZID,
                                   universe = entrezids_PR8_d60_all$ENTREZID,
                                   OrgDb = org.Mm.eg.db, 
                                   ont = "BP", 
                                   qvalueCutoff = 0.05, 
                                   readable = TRUE)
```

```{r}
cluster_summary_enrichment_PR8_d60 = data.frame(GO_enrichment_PR8_d60)
write.csv(cluster_summary_enrichment_PR8_d60, "../../Documents/machiels_lab_viral/output/DEA/clusterProfiler/PR8_d60/results_clusterProfiler_PR8_d60.csv")
```

```{r}
dotplot_goEnrichment_PR8_d60 = dotplot(GO_enrichment_PR8_d60, showCategory=50)
ggsave(filename = "dotplot_goEnrichment_PR8_d60.png", width = 45, height = 50, units = "cm", path =  "../../Documents/machiels_lab_viral/output/DEA/clusterProfiler/PR8_d60/")
```

```{r}
emapplot_goEnrichment_PR8_d60 = pairwise_termsim(GO_enrichment_PR8_d60)
emapplot_goEnrichment_PR8_d60 = emapplot(emapplot_goEnrichment_PR8_d60)


ggsave(filename = "emapplot_goEnrichment_PR8_d60.png", width = 45, height = 50, units = "cm", path =  "../../Documents/machiels_lab_viral/output/DEA/clusterProfiler/PR8_d60/")
```

```{r}
PR8_d60_logfold = markers_PR8_d60_filtered$avg_log2FC
names(PR8_d60_logfold) = markers_PR8_d60_filtered$gene

cnetplot(GO_enrichment_PR8_d60,
         categorySize="pvalue", 
         showCategory = 5, 
         foldChange=PR8_d60_logfold, 
         vertex.label.font=6)

ggsave(filename = "cnetplot_goEnrichment_PR8_d60.png", width = 45, height = 50, units = "cm", path =  "../../Documents/machiels_lab_viral/output/DEA/clusterProfiler/PR8_d60/")
```

## 2 C) Functional enrichment + pathway analysis of marker genes of PR8 d8 

```{r}
genes_PR8_d8 = markers_PR8_d8_filtered$gene
entrezids_PR8_d8_markers = select(org.Mm.eg.db, keys=genes_PR8_d8, keytype="SYMBOL", columns="ENTREZID")
entrezids_PR8_d8_markers = na.omit(entrezids_PR8_d8_markers)
```

```{r}
all_PR8_d8_genes = as.character(markers_PR8_d8$gene)
entrezids_PR8_d8_all = select(org.Mm.eg.db, keys=all_PR8_d8_genes, keytype="SYMBOL", columns="ENTREZID")
entrezids_PR8_d8_all = na.omit(entrezids_PR8_d8_all)
```

shows the relationship between the top 50 most significantly enriched GO terms (padj.), by grouping similar terms together. The color represents the p-values relative to the other displayed terms (brighter red is more significant) and the size of the terms represents the number of genes that are significant from our list.

```{r}
GO_enrichment_PR8_d8 = enrichGO(gene = entrezids_PR8_d8_markers$ENTREZID,
                                   universe = entrezids_PR8_d8_all$ENTREZID,
                                   OrgDb = org.Mm.eg.db, 
                                   ont = "BP", 
                                   qvalueCutoff = 0.05, 
                                   readable = TRUE)
```

```{r}
cluster_summary_enrichment_PR8_d8 = data.frame(GO_enrichment_PR8_d8)
write.csv(cluster_summary_enrichment_PR8_d8, "../../Documents/machiels_lab_viral/output/DEA/clusterProfiler/PR8_d8/results_clusterProfiler_PR8_d8.csv")
```

```{r}
dotplot_goEnrichment_PR8_d8 = dotplot(GO_enrichment_PR8_d8, showCategory=50)
ggsave(filename = "dotplot_goEnrichment_PR8_d8.png", width = 45, height = 50, units = "cm", path =  "../../Documents/machiels_lab_viral/output/DEA/clusterProfiler/PR8_d8/")
```

```{r}
emapplot_goEnrichment_PR8_d8 = pairwise_termsim(GO_enrichment_PR8_d8)
emapplot_goEnrichment_PR8_d8 = emapplot(emapplot_goEnrichment_PR8_d8)


ggsave(filename = "emapplot_goEnrichment_PR8_d8.png", width = 45, height = 50, units = "cm", path =  "../../Documents/machiels_lab_viral/output/DEA/clusterProfiler/PR8_d8/")
```

```{r}
PR8_d8_logfold = markers_PR8_d8_filtered$avg_log2FC
names(PR8_d8_logfold) = markers_PR8_d8_filtered$gene

cnetplot(GO_enrichment_PR8_d8,
         categorySize="pvalue", 
         showCategory = 5, 
         foldChange=PR8_d8_logfold, 
         vertex.label.font=6)

ggsave(filename = "cnetplot_goEnrichment_PR8_d8.png", width = 45, height = 50, units = "cm", path =  "../../Documents/machiels_lab_viral/output/DEA/clusterProfiler/PR8_d8/")
```
