---
title: "QC Experiment 1, day 60, lung"
format: 
  dashboard:
    orientation: columns
    

---
```{r setup}
#| warning: false
#| echo: false


library("easypackages")
libraries("Seurat", "tidyverse", "tidyseurat","ggExtra", "viridis", "ggridges","patchwork","Matrix", "ggpp","ggpubr","ggmosaic", "DT", "lemon"#, "gridExtra"
)

source(file="C:\\Users\\danne\\R_projects\\machiels_lab_viral\\dataset_QC_setup_chunk_exp1_lung.R")
#source(file="C:\\Users\\danne\\R_projects\\machiels_lab_viral\\dataset_QC_setup_chunk_exp1_bal.R")

#source(file="C:\\Users\\danne\\R_projects\\machiels_lab_viral\\dataset_QC_setup_chunk_exp2_bal.R")
#source(file="C:\\Users\\danne\\R_projects\\machiels_lab_viral\\dataset_QC_setup_chunk_exp2_lung.R")

```


# {.sidebar }


Thresholds applied to library:

|                       |                **lower**               |  **upper**                                           |
|-----------------------|----------------------------------------|------------------------------------------------------|
| **nCount quantile**   | `{r} lower_nCount_RNA_thresh_quantile` | `{r} upper_nCount_RNA_thresh_quantile`               |
| **nCount RNA**        | `{r} lower_nCount_RNA_thresh         ` | `{r} upper_nCount_RNA_thresh_non_scientific`               |


|                         |                **lower**               |  **upper**                                           |
|-------------------------|----------------------------------------|------------------------------------------------------|
| **quantile**            | none                                   |  `{r} upper_mito_thresh_quantile`                    |
| **mitochondrial genes** | none                                   |`{r} upper_mito_thresh`                               |        


|                       |                                        |  
|-----------------------|----------------------------------------|
| **deleted clusters** | `{r} deleted_clusters`                  |    


# global 

## Column {width=35%}

### Row {.tabset}


```{r}
#| title: UMAP
DimPlot(seurat_obj,label = T)+ggtitle("UMAP of first 10 Principal Components")

```
        
```{r}
#| title: UMAP filtered
#| 
filtered_seurat_obj <- seurat_obj |>
  filter(!(seurat_clusters %in% deleted_clusters)& 
           nCount_RNA<upper_nCount_RNA_thresh&
           nCount_RNA>lower_nCount_RNA_thresh&
           percent_mito<upper_mito_thresh&
           sampletag_multiplets=="single_hashtag")
filtered_seurat_obj |> DimPlot(label = T)


write_rds(filtered_seurat_obj, file = file_path_output
          )



```

### Row {.tabset}
```{r}
#| title: sample tags
seurat_obj |> count(sampletag_name) |>
  mutate(percent=round(100*(n/ncol(seurat_obj)),2)) |>
  ggplot(aes(sampletag_name, y=n, fill=sampletag_name)) +
  geom_col()+ 
  geom_text(aes(label=percent, x=sampletag_name, y=n))+
  theme_minimal()+
  theme_1+ ggtitle("sampletag_name")
```

```{r}
#| title: clusters

seurat_obj |> count(RNA_snn_res.0.5) |>
  mutate(percent=round(100*(n/ncol(seurat_obj)),2)) |>
  ggplot(aes(RNA_snn_res.0.5, y=n, fill=RNA_snn_res.0.5)) +
  geom_col()+ 
  geom_text(aes(label=percent, x=RNA_snn_res.0.5, y=n))+
  theme_minimal()+
  theme_1+
  ggtitle("seurat_slusters RNA_snn_res.0.5")

```


```{r}
#| title: Ms4a3

seurat_obj |> count(sampletag_Ms4a3) |>
  mutate(percent=round(100*(n/ncol(seurat_obj)),2)) |>
  ggplot(aes(sampletag_Ms4a3, y=n, fill=sampletag_Ms4a3)) +
  geom_col()+ 
  geom_text(aes(label=percent, x=sampletag_Ms4a3, y=n))+
  theme_minimal()+
  theme_1+
  ggtitle("seurat_slusters sampletag_Ms4a3")
```


```{r}
#| title: singlets
seurat_obj |> count(sampletag_multiplets) |>
  mutate(percent=round(100*(n/ncol(seurat_obj)),2)) |>
  ggplot(aes(sampletag_multiplets, y=n, fill=sampletag_multiplets)) +
  geom_col()+ 
  geom_text(aes(label=percent, x=sampletag_multiplets, y=n))+
  theme_minimal()+
  theme_1+
  ggtitle("seurat_slusters sampletag_multiplets")
```

## Column {width=40%}
### Row {.tabset}
```{r}
#| title: nCount marginal
multiplet_n_count_RNA_n_Feature<- seurat_obj |>
  ggplot(aes(nCount_RNA, nFeature_RNA, color=sampletag_multiplets)) +
  geom_point(size=0.5) +theme_bw() +xlim(c(-100,70000)) +ylim(c(-1000,8000)) +scale_color_manual(values = c( "grey","red","green"))+   geom_vline(aes(xintercept = lower_nCount_RNA_thresh),linetype = "dashed")+
  geom_vline(aes(xintercept = upper_nCount_RNA_thresh),linetype = "dashed")

ggMarginal(multiplet_n_count_RNA_n_Feature +
             theme(legend.position = "bottom") , type = "density", groupColour = TRUE, groupFill = TRUE,size = 3)  
```


```{r message=FALSE}
#| title: nCount marginal filtered clusters 1-7

multiplet_n_count_RNA_n_Feature_filtered<- filtered_seurat_obj |>mutate(seurat_clusters_num =as.numeric(as.character(seurat_clusters))) |>  filter(seurat_clusters_num %in% c(0:7))  |> ggplot(aes(nCount_RNA, nFeature_RNA, color=seurat_clusters)) +  
  geom_point(size=0.5) +
  theme_bw() +
  xlim(c(-100,70000)) +
  ylim(c(-1000,8000)) +
  #scale_color_manual(values = c( "grey","red","green"))  +  
  geom_vline(aes(xintercept = lower_nCount_RNA_thresh),linetype = "dashed")+    geom_vline(aes(xintercept = upper_nCount_RNA_thresh),linetype = "dashed")+ guides(colour = guide_legend(override.aes = list(size=7)))

ggMarginal(multiplet_n_count_RNA_n_Feature_filtered +theme(legend.position = "bottom") , type = "density", groupColour = TRUE, groupFill = TRUE,size = 3) 
max_num_cluster <-  filtered_seurat_obj|> mutate(seurat_clusters_num =as.numeric(as.character(seurat_clusters))) |> pull(seurat_clusters_num) |> max()
eval_filtered_15 <-  max_num_cluster>=15

```


```{r include=eval_filtered_15, message=FALSE}
#| title: nCount marginal filtered clusters 8-15

multiplet_n_count_RNA_n_Feature_filtered<- filtered_seurat_obj|> mutate(seurat_clusters_num =as.numeric(as.character(seurat_clusters))) |>  filter(seurat_clusters_num %in% c(8:15)) |>   ggplot(aes(nCount_RNA, nFeature_RNA, color=seurat_clusters)) + 
  geom_point(size=0.5) +
  theme_bw() +
  xlim(c(-100,70000)) +
  ylim(c(-1000,8000)) +
  #scale_color_manual(values = c( "grey","red","green"))  +  
  geom_vline(aes(xintercept = lower_nCount_RNA_thresh),linetype = "dashed")+    geom_vline(aes(xintercept = upper_nCount_RNA_thresh),linetype = "dashed")+ guides(colour = guide_legend(override.aes = list(size=7)))

ggMarginal(multiplet_n_count_RNA_n_Feature_filtered +theme(legend.position = "bottom") , type = "density", groupColour = TRUE, groupFill = TRUE,size = 3) 



max_num_cluster <-  filtered_seurat_obj|> mutate(seurat_clusters_num =as.numeric(as.character(seurat_clusters))) |> pull(seurat_clusters_num) |> max()
eval_filtered_15 <-  max_num_cluster>=15

```



```{r include=eval_filtered_15, message=FALSE, eval=eval_filtered_15}
#| title: nCount marginal filtered clusters >15
if (eval_filtered_15) {
  multiplet_n_count_RNA_n_Feature_filtered<- filtered_seurat_obj|> mutate(seurat_clusters_num =as.numeric(as.character(seurat_clusters))) |>  filter(seurat_clusters_num >15) |>   ggplot(aes(nCount_RNA, nFeature_RNA, color=seurat_clusters)) + 
  geom_point(size=0.5) +
  theme_bw() +
  xlim(c(-100,70000)) +
  ylim(c(-1000,8000)) +
  #scale_color_manual(values = c( "grey","red","green"))  +  
  geom_vline(aes(xintercept = lower_nCount_RNA_thresh),linetype = "dashed")+    geom_vline(aes(xintercept = upper_nCount_RNA_thresh),linetype = "dashed")+ guides(colour = guide_legend(override.aes = list(size=7)))

ggMarginal(multiplet_n_count_RNA_n_Feature_filtered +theme(legend.position = "bottom") , type = "density", groupColour = TRUE, groupFill = TRUE,size = 3)
  
  
}



```

```{r}
#| title: nCount boxplot unflitered

seurat_obj |> ggplot(aes(nCount_RNA, seurat_clusters, fill=seurat_clusters)) + geom_boxplot()+ theme_bw()

```

```{r}
#| title: nCount RNA boxplot filtered

filtered_seurat_obj |> ggplot(aes(nCount_RNA, seurat_clusters, fill=seurat_clusters)) + geom_boxplot()+ theme_bw()

```
```{r}
#| title: nCount RNA per condition unfiltered

seurat_obj |> ggplot(aes(nCount_RNA, sampletag_name, fill=sampletag_name)) + geom_boxplot()+ theme_bw()

```
```{r}
#| title: nCount RNA per condition filtered

sorted <- seurat_obj$sampletag_name |> unique() |> as.character()|> sort()

filtered_seurat_obj |> mutate(sampletag_name=factor(sampletag_name, levels=sorted  )) |> 
  ggplot(aes(nCount_RNA, sampletag_name, fill=sampletag_name)) +
  geom_boxplot()+ theme_bw()

```

```{r}
#| title: fraction multiplets, by nCOunt RNA quantile
 seurat_obj |>  
  mutate(ncount_bin=cut_number(nCount_RNA,50)) |> 
  separate(ncount_bin,sep = ",",into=c("throw","ncount_bin")) |>
  mutate(ncount_bin=as.numeric(str_remove(ncount_bin,"]"))) |> 
  group_by(ncount_bin) |>
  count(sampletag_multiplets) |> 
  mutate(percentage= 100*n/sum(n)) |> 
  ungroup() |> 
  ggplot(aes(ncount_bin, percentage, color=sampletag_multiplets)) + 
  geom_point(size=3) +
  theme_1+scale_color_manual(values = c( "grey","red","green"))+theme(legend.position = "right")

```

```{r}
#| title: fraction multiplets, by nCount RNA absolut
 seurat_obj |>  
  mutate(ncount_bin=cut_width(x=nCount_RNA,width=1000, boundary = 0)) |> 
  separate(ncount_bin,sep = ",",into=c("throw","ncount_bin")) |>
  mutate(ncount_bin=as.numeric(str_remove(ncount_bin,"]"))) |> 
  group_by(ncount_bin) |>
  count(sampletag_multiplets) |> 
  mutate(percentage= 100*n/sum(n)) |> 
  ungroup() |> 
  ggplot(aes(ncount_bin, percentage, color=sampletag_multiplets)) + 
  geom_point(size=3) +
  theme_1+
  theme_1+scale_color_manual(values = c( "grey","red","green"))
```



```{r}
#| title: mito

# 
plot_percent_mito_nFeature_RNA_dens <-  seurat_obj |> ggplot(aes(nCount_RNA, percent_mito)) + geom_density_2d() +theme_bw() +theme(legend.position = "none")+
        xlim(c(-1000,xlim_max)) +ylim(c(-0,ylim_max))

plot_percent_mito_nFeature_RNA_point <- seurat_obj |> ggplot(aes(nCount_RNA, percent_mito)) + geom_point(size=0.5) +theme_bw() +
        xlim(c(-1000,xlim_max)) +ylim(c(-0,6)) +
        geom_hline(aes(yintercept = upper_mito_thresh),linetype = "dashed")+
        geom_vline(aes(xintercept = lower_nCount_RNA_thresh),linetype = "dashed")+
  geom_vline(aes(xintercept = upper_nCount_RNA_thresh),linetype = "dashed")+
        geom_text(aes(xlim_max, upper_mito_thresh, label = as.character(upper_mito_thresh_quantile)), nudge_y =upper_mito_thresh/23)+
        geom_text(aes(lower_nCount_RNA_thresh,ylim_max, label = as.character(lower_nCount_RNA_thresh_quantile)), nudge_x =upper_nCount_RNA_thresh/23)+        geom_text(aes(upper_nCount_RNA_thresh,ylim_max, label = as.character(upper_nCount_RNA_thresh_quantile)), nudge_x =upper_nCount_RNA_thresh/20)

        
 plot_percent_mito_nFeature_RNA_dens

quantile_table <- tibble(
mito_quantiles=seurat_obj$percent_mito |>
        quantile(probs = (seq(0.93,0.99,by=0.01))),
nCount_quantiles=seurat_obj$nCount_RNA |>
        quantile(probs = (seq(0.93,0.99,by=0.01)))
)

quantile_table <- tibble(quantile_=c(
        seq(0.01,0.06,by=0.01),seq(0.93,0.99,by=0.01)) ) |> mutate(
        per_mito=round(seurat_obj$percent_mito |> quantile(probs = quantile_),2),
        nCount_RNA=round(seurat_obj$nCount_RNA |> quantile(probs = quantile_),2))


plot_percent_mito_nFeature_RNA_point <- plot_percent_mito_nFeature_RNA_point + annotate(geom = 'table',
           x=70000,
           y=6,
           size = 2,
           label=list(round(quantile_table, 2))) 


ggMarginal(plot_percent_mito_nFeature_RNA_point +
                   theme(legend.position = "none") , type = "density")
                   
```

```{r}
#| title: nCount density
multiplet_n_count_RNA_n_Feature_density<- seurat_obj |>  ggplot(aes(nCount_RNA, nFeature_RNA, color=sampletag_multiplets
                                                                    )) +geom_density_2d()+theme_bw() +xlim(c(-100,70000)) +ylim(c(-1000,8000)) +    geom_vline(aes(xintercept = lower_nCount_RNA_thresh),linetype = "dashed")+    geom_vline(aes(xintercept = upper_nCount_RNA_thresh),linetype = "dashed")

multiplet_n_count_RNA_n_Feature_density
```



### Row {.tabset}
```{r}
#| title: Elbow Plot
ElbowPlot(seurat_obj)
```

```{r}
#| title: PC 1/2
DimPlot(seurat_obj, reduction = "pca")
```
```{r}
#| title: PC  3/4
DimPlot(seurat_obj, reduction = "pca", dims = c(3,4))
```
```{r}
#| title: PC  4/5
DimPlot(seurat_obj, reduction = "pca", dims = c(4,5))
```
```{r}
#| title: PC  6/7
DimPlot(seurat_obj, reduction = "pca", dims = c(6,7))
```
```{r}
#| title: PC  8/9
DimPlot(seurat_obj, reduction = "pca", dims = c(8,9))
```

## Column {.tabset}
```{r}
#| fig-width: 12
#| fig-height: 12
#| echo: false
#| title: all cells

seurat_obj <- seurat_obj |>
  mutate(rna_group=ifelse(nCount_RNA<lower_nCount_RNA_thresh,"low_RNA",
                                    ifelse(nCount_RNA>upper_nCount_RNA_thresh,"high_RNA",
                                           "norm_RNA"
                                    )),
                     mito_group=ifelse(percent_mito>upper_mito_thresh,"high_mito","norm_mito")) |> 
  mutate(rna_mito=as_factor(paste(rna_group,mito_group, sep="..."))) #|> 

seurat_obj |> 
  ggplot()+
  geom_mosaic(aes(x=product(rna_mito,sampletag_multiplets),fill=rna_mito, alpha=sampletag_multiplets))+theme_1+theme(axis.text  = element_blank())+
  theme(legend.position = "right",axis.title = element_text(size=22),
        legend.text=element_text(size=22),
        axis.text = element_text(size = 22),
        axis.text.y = element_blank()
        )+
  scale_fill_discrete(guide= guide_legend(reverse=TRUE))

```


```{r}
#| fig-width: 7
#| fig-height: 15
#| title: seurat clusters
#| warning: false
cluster_mosaic_plot <- seurat_obj  |> 
        ggplot()+
        geom_mosaic(aes(x=product(rna_mito,sampletag_multiplets),
                        fill=rna_mito, alpha=sampletag_multiplets))+
        theme_1+
        theme(axis.text  = element_blank(),
              axis.text.x = element_blank())+facet_wrap(~seurat_clusters,ncol = 3)#+ theme(strip.text = element_text(size=22))

cluster_mosaic_plot

```

```{r}
#| title: cell cycle


# A list of cell cycle markers, from Tirosh et al, 2015, is loaded with Seurat.  We can
# segregate this list into markers of G2/M phase and markers of S phase
s.genes <- cc.genes$s.genes |> str_to_title()
g2m.genes <- cc.genes$g2m.genes|> str_to_title()

seurat_obj<- CellCycleScoring(seurat_obj, s.features = s.genes, g2m.features = g2m.genes)

DimPlot(seurat_obj, group.by = "Phase")

```
```{r}
#| title: cell cycle
cluster_cell_cycle_mosaic_plot_all <- seurat_obj  |> 
        ggplot()+
        geom_mosaic(aes(x=product(Phase,sampletag_multiplets),
                        fill=Phase, alpha=sampletag_multiplets))+
        theme_1+
        theme(axis.text  = element_blank(),
              axis.text.x = element_blank(),
              legend.position = "bottom")
cluster_cell_cycle_mosaic_plot_all 
```

```{r}
#| title: cell cycle
#| fig-width: 7
#| fig-height: 15
#| warning: false
cluster_cell_cycle_mosaic_plot <- seurat_obj  |> 
        ggplot()+
        geom_mosaic(aes(x=product(Phase,sampletag_multiplets),
                        fill=Phase, alpha=sampletag_multiplets))+
        theme_1+
        theme(axis.text  = element_blank(),
              axis.text.x = element_blank())+facet_wrap(~seurat_clusters,ncol = 3)
cluster_cell_cycle_mosaic_plot 
```

```{r}
#| fig-width: 7
#| fig-height: 15
#| title: condition
#| warning: false

seurat_obj  |> 
        ggplot()+
        geom_mosaic(aes(x=product(rna_mito,sampletag_multiplets),
                        fill=rna_mito, alpha=sampletag_multiplets))+
        theme_1+
        theme(axis.text  = element_blank(),
              axis.text.x = element_blank())+
                      facet_wrap(~sampletag_name,
                                 ncol = 3) + theme(strip.text = element_text(size=22))
```
# gene-level

## Column 

### Row {,.tabset}
```{r}
seurat_obj <- BuildClusterTree(seurat_obj)
PlotClusterTree(seurat_obj, direction = "downwards")
```

### Row {}
```{r}
#| title: Data

de_genes_tbl |> mutate(cluster=as_factor(cluster)) |>
  relocate(cluster) |> 
  datatable(
    colnames = c("cluster ","gene","pval", "avg_log2FC", "pct.1", "pct.2","p_val_adj"),
    filter = 'top',
    options = list(dom = 'tp',
     paging = TRUE,
    #   lengthMenu = c(5, 10, 25),
    #   pageLength = 10, 
      scrollY = TRUE
     ),
  )
```
## Column 

### Row{.tabset}
```{r}
#| warning: false
#| echo: false
#load packages

feature_umap <- function(genes, title_="", caption_=""){
  
        seurat_obj |> join_features(genes) |> mutate(.feature=factor(.feature,levels=genes)) |> ggplot(aes(umap_1,umap_2,color=.abundance_RNA))+geom_point() +
  scale_color_viridis_c()+ theme_void()+coord_fixed()+
  facet_wrap(~.feature, ncol = 4) +
    labs(title =title_,
         caption =caption_ )+
    theme(plot.caption = element_text(size=6, face="bold"))
}

clus_0_genes <-  c(de_genes_tbl |> filter(cluster==0) |> slice_max(avg_log2FC,n = 5) |> pull(gene), c("H2-Ab1"))
title_feature_umap_clus_0 <- "Macrophage, MHC complex"
caption_feature_umap_clus_0 <-"" 
```


```{r, cluster_0_feature_umap, eval= eval_0, include=eval_0}
#| title: cluster 0
#| fig-height: 10
feature_umap(clus_0_genes, title_= title_feature_umap_clus_0, caption_ = caption_feature_umap_clus_1) 
```

```{r, cluster_1_feature_umap, eval= eval_1, include=eval_1}
#| title: cluster 1
feature_umap(genes=clus_1_genes,title_feature_umap_clus_1, caption=caption_feature_umap_clus_1)
```

```{r, cluster_2_feature_umap, eval= eval_2, include=eval_2}
#| title: cluster 2
feature_umap(clus_2_genes,title_feature_umap_clus_2, caption=caption_feature_umap_clus_2)
```

```{r, cluster_3_feature_umap, eval= eval_3, include=eval_3}
#| title: cluster 3
feature_umap(clus_3_genes,title_feature_umap_clus_3, caption=caption_feature_umap_clus_3)
```

```{r, cluster_4_feature_umap, eval= eval_4, include=eval_4}
#| title: cluster 4
feature_umap(clus_4_genes,title_feature_umap_clus_4, caption=caption_feature_umap_clus_4)
```

```{r, cluster_5_feature_umap, eval= eval_5, include=eval_5}
#| title: cluster 5

feature_umap(clus_5_genes,title_feature_umap_clus_5, caption=caption_feature_umap_clus_5)
```

```{r, cluster_6_feature_umap, eval= eval_6, include=eval_6}
#| title: cluster 6
feature_umap(clus_6_genes,title_feature_umap_clus_6, caption=caption_feature_umap_clus_6)
```

```{r, cluster_7_feature_umap, eval= eval_7, include=eval_7}
#| title: cluster 7

feature_umap(clus_7_genes,title_feature_umap_clus_7, caption=caption_feature_umap_clus_7)
```
```{r, cluster_8_feature_umap, eval= eval_8, include=eval_8}
#| title: cluster 8
feature_umap(clus_8_genes,title_feature_umap_clus_8, caption=caption_feature_umap_clus_8)

```
```{r, cluster_9_feature_umap, eval= eval_9, include=eval_9}
#| title: cluster 9

feature_umap(clus_9_genes,title_feature_umap_clus_9, caption=caption_feature_umap_clus_9)
```
```{r, cluster_10_feature_umap, eval= eval_10, include=eval_10}
#| title: cluster 10

feature_umap(clus_10_genes,title_feature_umap_clus_10, caption=caption_feature_umap_clus_10)
```

```{r, cluster_10_extra, eval= eval_10_extra, include= eval_10_extra}
#| title: cluster 10 mt graph

clus=10
seurat_obj |> mutate(cluster_10=ifelse(RNA_snn_res.0.5==clus, paste("cluster", as.character(clus)), "other"))|> 
  arrange(desc(cluster_10)) |> 
  ggplot(aes(nCount_RNA,percent_mito, color=cluster_10)) +geom_point()
```


```{r, cluster_11_feature_umap, eval= eval_11, include=eval_11}
#| title: cluster 11

feature_umap(clus_11_genes,title_feature_umap_clus_11, caption=caption_feature_umap_clus_11)

```

```{r, cluster_12_feature_umap, eval= eval_12, include=eval_12}
#| title: cluster 12
#| 
feature_umap(clus_12_genes,title_feature_umap_clus_12, caption=caption_feature_umap_clus_12)

```
```{r, cluster_13_feature_umap, eval= eval_13, include=eval_13}
#| title: cluster 13

feature_umap(clus_13_genes,title_feature_umap_clus_13, caption=caption_feature_umap_clus_13)

```

```{r, cluster_14_feature_umap, eval= eval_14, include=eval_14}
#| title: cluster 14

feature_umap(clus_14_genes,title_feature_umap_clus_14, caption=caption_feature_umap_clus_14)
```

```{r, cluster_15_feature_umap, eval= eval_15, include=eval_15}
#| title: cluster 15

feature_umap(clus_15_genes,title_feature_umap_clus_15, caption=caption_feature_umap_clus_15)
```




```{r, cluster_16_feature_umap, eval= eval_16, include=eval_16}
#| title: cluster 16

feature_umap(clus_16_genes,title_feature_umap_clus_16, caption=caption_feature_umap_clus_16)
```

```{r, cluster_16_extra, eval= eval_16_extra}
#| title: cluster 16 expr
#| 
seurat_obj |> mutate(cluster_highlight=ifelse(RNA_snn_res.0.5==16, "cluster 16", "other")) |> 
  join_features(clus_16_genes) |> mutate(.feature=factor(.feature, levels=clus_16_genes)) |> 
  ggplot(aes(cluster_highlight,.abundance_RNA), color=cluster_highlight) + 
  geom_violin()+geom_jitter()+
  theme_minimal()+
  facet_wrap(~.feature)

```


```{r, cluster_17_feature_umap, eval= eval_17, include=eval_17}
#| title: cluster 17

feature_umap(clus_17_genes,title_feature_umap_clus_17, caption=caption_feature_umap_clus_17)
```
```{r, cluster_18_feature_umap, eval= eval_18, include=eval_18}
#| title: cluster 18

feature_umap(clus_18_genes,title_feature_umap_clus_18, caption=caption_feature_umap_clus_18)

```

```{r, cluster_19_feature_umap, eval= eval_19, include=eval_19}
#| title: cluster 19

feature_umap(clus_19_genes,title_feature_umap_clus_19, caption=caption_feature_umap_clus_19)
```

```{r, cluster_20_feature_umap, eval= eval_20, include=eval_20}
#| title: cluster 20

feature_umap(clus_20_genes,title_feature_umap_clus_20, caption=caption_feature_umap_clus_20)
```

```{r, cluster_21_feature_umap, eval= eval_21, include=eval_21}
#| title: cluster 21

feature_umap(clus_21_genes,title_feature_umap_clus_21, caption=caption_feature_umap_clus_21)
```
### Row {.tabset}
```{r}
#| title: cluster 0

clus_highlight <- function(clus){
  seurat_obj |> mutate(cluster_highlight=ifelse(RNA_snn_res.0.5==clus, paste("cluster", as.character(clus)), "other")) |>
  ggplot(aes(umap_1,umap_2,color=cluster_highlight ))+geom_point() +theme_minimal()+ theme(legend.position = "none")
  
}

clus_highlight(0)
```
```{r}
#| title: cluster 1

clus_highlight(1)

```

```{r}
#| title: cluster 2

clus_highlight(2)

```
```{r}
#| title: cluster 3

clus_highlight(3)

```
```{r}
#| title: cluster 4

clus_highlight(4)

```
```{r}
#| title: cluster 5

clus_highlight(5)

```
```{r}
#| title: cluster 6

clus_highlight(6)

```
```{r}
#| title: cluster 7

clus_highlight(7)

```

```{r}
#| title: cluster 8

clus_highlight(8)

```
```{r}
#| title: cluster 9

clus_highlight(9)

```

```{r}
#| title: cluster 10

clus_highlight(10)

```

```{r}
#| title: cluster 11

clus_highlight(11)

```




```{r}
#| title: cluster 12

clus_highlight(12)

```

```{r}
#| title: cluster 13

clus_highlight(13)

```

```{r}
#| title: cluster 14

clus_highlight(14)

```

```{r}
#| title: cluster 15

clus_highlight(15)

```

```{r}
#| title: cluster 16

clus_highlight(16)

```

```{r}
#| title: cluster 17

clus_highlight(17)

```


```{r}
#| title: cluster 17
clus_highlight(17)
```
```{r}
#| title: cluster 18
clus_highlight(18)
```
```{r}
#| title: cluster 19
clus_highlight(19)
```
```{r}
#| title: cluster 20
clus_highlight(20)
```
```{r}
#| title: cluster 21

clus_highlight(21)

```






