```{r}

```

---
title: "QC Experiment 1, day 60, bal"
format: 
  html:
    code-fold: true
editor: visual
toc: true
execute:
  warning: false
  message: false
---

# DANIEL: 13372 from 18000

# 1 Load in data, packages + initialize functions

```{r, echo=FALSE}
library("easypackages")
set.seed(2023)
libraries("tidyverse", "Seurat", "tidyseurat", "tibble", 'scCustomize', 'RCurl', 'ensembldb', 'AnnotationHub')

```

```{r}
file_path <- vector("list") 
file_path$output <- "../../Documents/machiels_lab_viral/output"
file_path$intermediate_data<- "../../Documents/machiels_lab_viral/intermediate_data/"
file_path$raw_data_root <- "../../Documents/machiels_lab_viral/raw_data/machiels_lab/viral/"
file_names_tibble <- tribble(
    ~library_name,
    ~raw_seurat_obj_path,
    ~path_sample_tag_calls,
    ~write_processed_seurat_obj_path,

    "bal",
    #library_name
    paste0(file_path$intermediate_data,"seurat_obj_experiment_1_combined_bal_raw_dbec.rds"),
    #raw_seurat_obj_path
    paste0(file_path$raw_data_root,"2023-10-02_output_bal/Output_BAL/","BD-Analysis-BMachiels_Sample_Tag_Calls.csv"),
    #path_sample_tag_calls
    paste0(file_path$intermediate_data,"seurat_obj_experiment_1_bal_workflowed.rds"),
    #write_processed_seurat_obj_path
)
```

this function initalizes the sampletag_name metadata column in the seurat object at hand. If the sample (an individual cell) corresponds to a 'singlet' the sample will be annotated according to its condition (virustype or mock) and wether it is Ms4a3 positive or negative.

```{r}
seurat_sampletag_name <- function(seurat_obj, sample_tag_calls){
    seurat_obj$sampletag_name <- sample_tag_calls  |>
        right_join(tibble(Cell_Index=colnames(seurat_obj))) |> 
        mutate(Sample_Name=str_replace(Sample_Name,pattern="\\-",replacement = "_"))|>
        mutate(Sample_Name=str_replace_all(Sample_Name,pattern="\\+",
                                           replacement = "_pos")) |>
        mutate(Sample_Name=str_replace_all(Sample_Name,pattern="\\-", 
                                           replacement = "_neg")) |>
        pull(Sample_Name) |>  as_factor()
    seurat_obj <- seurat_obj |>  mutate(sampletag_multiplets=case_when(
        sampletag_name=="Multiplet" ~ "multiplet",
        sampletag_name=="Undetermined" ~"undeterminded",
        TRUE ~ "single_hashtag")) |> 
        mutate(sampletag_multiplets=factor(sampletag_multiplets,
                                           levels=c("undeterminded",
                                                    "multiplet",
                                                    "single_hashtag"))) 
    #create separate groups based on Ms4a3
    seurat_obj <- seurat_obj |>  mutate(sampletag_Ms4a3=case_when(
        sampletag_name=="Multiplet" ~ "multiplet",
        sampletag_name=="Undetermined" ~"undeterminded",
        str_detect(sampletag_name, pattern="pos") ~ "Ms4a3_pos",
        str_detect(sampletag_name, pattern="neg") ~ "Ms4a3_neg",
        TRUE ~ "single_hashtag"))|> 
        mutate(sampletag_Ms4a3=as_factor(sampletag_Ms4a3))
    
    seurat_obj$condition <- seurat_obj |> pull(sampletag_name) |> 
      str_split_i(i=1, pattern = "_") |> as_factor()
    seurat_obj$virus <- tibble(sampletag_name=seurat_obj |> pull(sampletag_name)) |>       
    separate(sampletag_name,into = c("virus")) |> pull("virus")
    #sort the factors in their desired format
    seurat_obj@meta.data$sampletag_Ms4a3 = factor(seurat_obj@meta.data$sampletag_Ms4a3,     levels = c("Ms4a3_pos", "Ms4a3_neg"))
    seurat_obj@meta.data$sampletag_name = factor(seurat_obj@meta.data$sampletag_name,     levels = c("Multiplet", "Undetermined", "Mock_Ms4a3_neg", "Mock_Ms4a3_pos",                      "MuHV4_Ms4a3_neg", "MuHV4_Ms4a3_pos", "MAV1_Ms4a3_neg",                               "MAV1_Ms4a3_pos", "PR8_Ms4a3_neg", "PR8_Ms4a3_pos", "PVM_Ms4a3_neg",                  "PVM_Ms4a3_pos"))
    
    return(seurat_obj)
}
```

```{r, echo=FALSE}
seurat_d60_bal_raw = read_rds(file_names_tibble$raw_seurat_obj_path)
for (line in 1:nrow(file_names_tibble)) {
sample_tag_calls <- read_csv(file_names_tibble[[line,"path_sample_tag_calls"]], skip = 7) |>
    mutate(Cell_Index=as.character(Cell_Index))
}
```

# 2 QC basic metrics"../../Documents/machiels_lab_viral/Case_study_annotation/featureplots"

to start off the raw data contains 27774 genes expressed among 14000 different cells.

Many of those might contain a high degree of mitochondrial genes indicating that the individual cells or damaged and/or are dying. Because when the cell membrane breaks down mRNA from the cytoplasm starts to leak out leaving out a higher degree of mitochondrial RNA which is more resistant to degradation.

Many of those might be so called doulets. When two cells are accidentally captured together during the isolation process, their combined gene expression profile is recorded. This creates a "doublet" event, where the resulting data represents a mixture of two distinct cell types. This skews the overall data and introduces misleading information about the true expression patterns within the individual cell populations.

Another subpopulation of single cell samples that need to be filtered out are the so called undetermined single cell samples. These single cell samples might not be a single cell but a cellular fragment, debris, or even an artifact. These "undetermined" events contribute noise to the data and can hinder the accurate identification of cell types.

nCount RNA = #UniqueMolecularIdentifiers / Cell = amountUMI

nFeature RNA = #genes detected / cell =nGenes

```{r, echo=FALSE}

seurat_d60_bal_raw = seurat_sampletag_name(seurat_obj = seurat_d60_bal_raw, sample_tag_calls = sample_tag_calls)
seurat_d60_bal_raw$amountUMI = seurat_d60_bal_raw$nCount_RNA
seurat_d60_bal_raw$amountGenes = seurat_d60_bal_raw$nFeature_RNA
```

## 2.1 Calculate the complexity (amount genes detected per UMI)

the amount of genes detected per UMI aka the amount of genes detected per cell

a higher degree corresponds to a more diverse and complex transcriptome within the individual cell

a lower degree corresponds to cells suferring from technical issues during manipulation

a lower degree corresponds to cells of poor quality like stressed damaged or dying cells

```{r, echo=FALSE}
seurat_d60_bal_raw$log10GenesPerUMI = log10(seurat_d60_bal_raw$nFeature_RNA) / log10(seurat_d60_bal_raw$nCount_RNA)

```

## 2.2 Calculate the mitochondrial ratio

```{r, echo=FALSE}
seurat_d60_bal_raw$mitoRatio <- PercentageFeatureSet(object = seurat_d60_bal_raw, pattern = "^mt-")
seurat_d60_bal_raw$mitoRatio <- seurat_d60_bal_raw@meta.data$mitoRatio / 100
```

## 

# 3 Assessment and filtering of the general QC metrics

## 3.1 Remove doublets and undetermined cells

Make sure the seurat object only contains singlets i.e. filtering doublets and undetermined cells out

```{r, echo=FALSE}
seurat_d60_bal_raw = dplyr::filter(seurat_d60_bal_raw, sampletag_multiplets == "single_hashtag")
seurat_d60_bal_raw
```

### after filtering for singlets only 15593 cells out of 18000 were retained

## 3.2 Visualize the distribution of the amount of cells amongst the conditions

```{r}
#| fig-width: 10
#| fig-height: 8
seurat_d60_bal_raw %>% 
  	ggplot(aes(x=condition, fill=condition)) + 
  	geom_bar() +
    scale_y_continuous(breaks=c(0,500,1000,1500,2000,2500,3000,3500)) + 
  	theme_classic() +
  	theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  	theme(plot.title = element_text(hjust=0.5, face="bold")) +
  	ggtitle("Amount cells in the different conditions")
```

Before doing any subsequent filtering we see a quite substantial higher amount of cells corresponding to the viral conditions and about an even distribution of amount of cells betweeen the viral conditions

## 3.2 Visualize the distribution of the amount of cells amongst the conditions and Ms4a3

```{r}
#| fig-width: 10
#| fig-height: 8

seurat_d60_bal_raw %>% 
  	ggplot(aes(x=sampletag_name, fill=sampletag_name)) + 
  	geom_bar(stat = "count") +
    scale_y_continuous(breaks=c(0,500,1000,1500,2000,2500,3000,3500)) +
  	theme_classic() +
  	theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  	theme(plot.title = element_text(hjust=0.5, face="bold")) +
  	ggtitle("Amount cells in the different conditions and Ms4a3")
```

## 3.3 Visualize the distribution of the amount of cells amongst Ms4a3

```{r}
#| fig-width: 10
#| fig-height: 8
seurat_d60_bal_raw %>% 
  	ggplot(aes(x=sampletag_Ms4a3, fill=sampletag_Ms4a3)) + 
  	geom_bar() +
  	theme_classic() +
  	theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  	theme(plot.title = element_text(hjust=0.5, face="bold")) +
  	ggtitle("Amount cells in the different conditions for Ms4a3")
```

### More Ms4a3 negative cells compared to Ms4a3 positive. apparently a lot more PVM Ms4a3- cells compared to the rest

## 3.4 Density and plot showing amount of UMI per cell amongst the different conditions

```{r}
#| fig-width: 10
#| fig-height: 8
seurat_d60_bal_raw %>% 
  	ggplot(aes(color=condition, x=amountUMI, fill= condition)) + 
  	geom_density(alpha = 0.2) + 
  	theme_classic() +
  	ylab("Cell density") +
  	geom_vline(xintercept = 5000)
```

## 3.5 Density and plot showing amount of UMI per cell amongst the different conditions and Ms4a3 status

```{r}
#| fig-width: 10
#| fig-height: 8
seurat_d60_bal_raw %>% 
  	ggplot(aes(color=sampletag_name, x=amountUMI, fill= sampletag_name)) + 
  	geom_density(alpha = 0.2) + 
  	theme_classic() +
  	ylab("Cell density") +
  	geom_vline(xintercept = 500)
```

## 3.6 Density and plot showing amount of UMI per cell amongst Ms4a3 status

```{r}
#| fig-width: 10
#| fig-height: 8
seurat_d60_bal_raw %>% 
  	ggplot(aes(color=sampletag_Ms4a3, x=amountUMI, fill= sampletag_Ms4a3)) + 
  	geom_density(alpha = 0.2) + 
  	theme_classic() +
  	ylab("Cell density") +
  	geom_vline(xintercept = 500)
```

### Sequencing depth is appropriate generally when the UMI counts per cell are below 500 they cannot be used or when between 500-1000 it can be used but sequencing should be deeper. In this case it looks appropriate

## 3.7 Density and plot showing amount of genes per cell amongst the conditions

```{r}
#| fig-width: 10
#| fig-height: 8
seurat_d60_bal_raw %>% 
  	ggplot(aes(color=condition, x=amountGenes, fill= condition)) + 
  	geom_density(alpha = 0.2) + 
  	theme_classic() +
  	ylab("Cell density") 
```

## 3.8 Density and plot showing amount of genes per cell amongst the conditions and Ms4a3

```{r}
#| fig-width: 10
#| fig-height: 8
seurat_d60_bal_raw %>% 
  	ggplot(aes(color=sampletag_name, x=amountGenes, fill= sampletag_name)) + 
  	geom_density(alpha = 0.2) + 
  	theme_classic() +
  	ylab("Cell density") 

```

## 3.9 Density and plot showing amount of genes per cell amongst Ms4a3

```{r}
#| fig-width: 10
#| fig-height: 8
seurat_d60_bal_raw %>% 
  	ggplot(aes(color=sampletag_Ms4a3, x=amountGenes, fill= sampletag_Ms4a3)) + 
  	geom_density(alpha = 0.2) + 
  	theme_classic() +
  	ylab("Cell density") 
```

## 3.10 number of genes versus the number of UMIs coloured by the proportion of mitochondrial reads

Visualize the correlation between genes detected and number of UMIs and determine whether strong presence of cells with low numbers of genes/UMIsVisualize the correlation between genes detected and number of UMIs and determine whether strong presence of cells with low numbers of genes/UMIs.

From this we see only cells in the bottom left, these cells are likely of bad quality since they have a lower amount of UMI and genes. Combine this with the observation that they also contain a higher fraction of mitochondrial reads it can be fair to assume that these cells are possiby dying or are of bad quality.

#### Although increased circulating mitochondrial DNA (mtDNA) has been reported in chronic lung diseases (<https://www.researchgate.net/publication/333656561_PINK1_attenuates_mtDNA_release_in_alveolar_epithelial_cells_and_TLR9_mediated_profibrotic_responses>)

```{r}
#| fig-width: 10
#| fig-height: 10
seurat_d60_bal_raw %>% 
  	ggplot(aes(x=amountUMI, y=amountGenes, color=mitoRatio)) + 
  	geom_point() + 
	  scale_colour_gradient(low = "green", high = "red") +
  	stat_smooth(method=lm) +
  	scale_x_log10() + 
  	scale_y_log10() + 
  	theme_classic() +
  	facet_wrap(~sampletag_name, scales = "free_x") +
    theme(axis.text.x = element_text(size = 8),
          axis.text.y = element_text(size = 8),
          legend.text = element_text(size = 14),
          legend.title = element_text(size = 14),
          plot.title = element_text(size = 14, hjust = 0.5))
```

## 3.11 Boxplot investigating distribution of genes and UMI detected per cell

```{r}
#| fig-width: 10
#| fig-height: 10
seurat_d60_bal_raw %>%
  ggplot(aes(x = sampletag_name, y = amountGenes, fill = sampletag_name)) +
  geom_boxplot(outlier.shape = "triangle") +  # Hide outliers initially
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) +
  ggtitle("Amount Cells vs. Amount Genes")
```

## 3.12 Boxplot investigating distribution of UMI detected per cell

```{r}
#| fig-width: 10
#| fig-height: 10
seurat_d60_bal_raw %>% 
  	ggplot(aes(x=sampletag_name, y=amountUMI, fill=sampletag_name)) + 
  	geom_boxplot(outlier.shape = "triangle") + 
  	theme_classic() +
  	theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  	theme(plot.title = element_text(hjust=0.5, face="bold")) +
  	ggtitle("amount cells vs amount UMI")
```

## 3.13 Density and plot showing complexity amongst the conditions

```{r}
#| fig-width: 10
#| fig-height: 8
seurat_d60_bal_raw %>% 
  ggplot(aes(color=condition, x=log10GenesPerUMI, fill= condition)) + 
  geom_density(alpha = 0.2) + 
  theme_classic()
```

## 3.13 Density and plot showing complexity amongst the conditions and Ms4a3

```{r}
#| fig-width: 10
#| fig-height: 8
seurat_d60_bal_raw %>% 
  ggplot(aes(color=sampletag_name, x=log10GenesPerUMI, fill= sampletag_name)) + 
  geom_density(alpha = 0.2) + 
  theme_classic()
```

## 3.14 Density and plot showing complexity amongst Ms4a3

```{r}
#| fig-width: 10
#| fig-height: 8
seurat_d60_bal_raw %>% 
  ggplot(aes(color=sampletag_Ms4a3, x=log10GenesPerUMI, fill= sampletag_Ms4a3)) + 
  geom_density(alpha = 0.2) + 
  theme_classic()
```

## 3.15 Density plot showing rate of mitochondrial transcripts amongst conditions

```{r}
#| fig-width: 10
#| fig-height: 8
seurat_d60_bal_raw %>% 
  ggplot(aes(color=condition, x=mitoRatio, fill= condition)) + 
  geom_density(alpha = 0.2) +
  theme_classic() 
```

## 3.16 Density plot showing rate of mitochondrial transcripts amongst conditions and Ms4a3

```{r}
#| fig-width: 10
#| fig-height: 8
seurat_d60_bal_raw %>% 
  ggplot(aes(color=sampletag_name, x=mitoRatio, fill= sampletag_name)) + 
  geom_density(alpha = 0.2) +
  theme_classic() 
  
```

## 3.17 Density plot showing rate of mitochondrial transcripts amongst Ms4a3

```{r}
#| fig-width: 10
#| fig-height: 8
seurat_d60_bal_raw %>% 
  ggplot(aes(color=sampletag_Ms4a3, x=mitoRatio, fill= sampletag_Ms4a3)) + 
  geom_density(alpha = 0.2) +
  theme_classic() 
```

# 4 Filtering of poor quality cells

```{r}
seurat_d60_bal_filtered = dplyr::filter(seurat_d60_bal_raw, 
                                (mitoRatio < 0.2) &
                                (amountGenes > 500) &
                                (amountUMI > 5000) &
                                (log10GenesPerUMI > 0.8))

```

# 14228 out of 18000 cells were retained

## 4.1 Visualize the distribution of the amount of cells amongst the conditions after cell level filtering

```{r}
#| fig-width: 10
#| fig-height: 8
seurat_d60_bal_filtered %>% 
  	ggplot(aes(x=condition, fill=condition)) + 
  	geom_bar() +
    scale_y_continuous(breaks=c(0,500,1000,1500,2000,2500,3000,3500)) + 
  	theme_classic() +
  	theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  	theme(plot.title = element_text(hjust=0.5, face="bold")) +
  	ggtitle("Amount cells in the different conditions")
```

## 4.2 Visualize the distribution of the amount of cells amongst the conditions and Ms4a3 after cell level filtering

```{r}
#| fig-width: 10
#| fig-height: 8

seurat_d60_bal_filtered %>% 
  	ggplot(aes(x=sampletag_name, fill=sampletag_name)) + 
  	geom_bar(stat = "count") +
    scale_y_continuous(breaks=c(0,500,1000,1500,2000,2500,3000,3500)) +
  	theme_classic() +
  	theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  	theme(plot.title = element_text(hjust=0.5, face="bold")) +
  	ggtitle("Amount cells in the different conditions and Ms4a3")
```

## 4.3 Visualize the distribution of the amount of cells amongst Ms4a3 after cell level filtering

```{r}
seurat_d60_bal_filtered %>% 
  	ggplot(aes(x=sampletag_Ms4a3, fill=sampletag_Ms4a3)) + 
  	geom_bar(stat = "count") +
  	theme_classic() +
  	theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  	theme(plot.title = element_text(hjust=0.5, face="bold")) +
  	ggtitle("Amount cells in the different conditions for Ms4a3")
```

# 5 Gene - level filtering

We do not want to retain genes that are only expressed exclusively in a very low amount of cells as well

```{r}
countsDataset <- GetAssayData(object = seurat_d60_bal_filtered, slot = "counts")

nonzero <- countsDataset > 0
# only retain genes who are expressed in 5 or more cells so we retain rare possible muts
keep_genes <- Matrix::rowSums(nonzero) >= 5

# Only keeping those genes expressed in more than 10 cells
filtered_counts <- countsDataset[keep_genes, ]

#create a new object filtered1 because otherwise you lose the other assay data
seurat_d60_bal_filtered1 <- CreateSeuratObject(counts = filtered_counts, meta.data = seurat_d60_bal_filtered@meta.data)
# copy them back from the object where the genes werent filtered
seurat_d60_bal_filtered1@assays$protein = seurat_d60_bal_filtered@assays$protein
seurat_d60_bal_filtered1@assays$sampletags = seurat_d60_bal_filtered@assays$sampletags
```

# After QC we retain 21559 genes and 14228 cells from 27314 genes and 18000 cells

# 6 Normalization

## 6.1 Normalize the counts

```{r}
seurat_d60_bal_filtered_normalized = NormalizeData(seurat_d60_bal_filtered1)
```

## 6.2 Check for effect of cell cycle

if cells are differentiating it is expected that there are clear clustering differences between G2M and S.

### 6.2.1 Get cell cycle genes for mouse

```{r}
cell_cycle_file <- getURL("https://raw.githubusercontent.com/hbc/tinyatlas/master/cell_cycle/Mus_musculus.csv") 
cell_cycle_genes <- read.csv(text = cell_cycle_file)

```

### 6.2.2 convert the ensembl ID geneformat to geneID

The genes from this list are all in ENSEMBL format while those in the seurat object are the gene ID's =\> convert it

```{r, echo=FALSE}
h <- AnnotationHub()

# Access the Ensembl database for organism
ahDb <- query(h, 
              pattern = c("Mus musculus", "EnsDb"), 
              ignore.case = TRUE)

# Acquire the latest annotation files
id <- ahDb %>%
        mcols() %>%
        rownames() %>%
        tail(n = 1)

# Download the appropriate Ensembldb database
edb <- h[[id]]

# Extract gene-level information from database
annotations <- genes(edb, 
                     return.type = "data.frame")

# Select annotations of interest
annotations <- annotations %>%
        dplyr::select(gene_id, gene_name, seq_name, gene_biotype, description)

# Get gene names for Ensembl IDs for each gene
cell_cycle_markers <- dplyr::left_join(cell_cycle_genes, annotations, by = c("geneID" = "gene_id"))

# Acquire the S phase genes
s_genes <- cell_cycle_markers %>%
        dplyr::filter(phase == "S") %>%
        pull("gene_name")
        
# Acquire the G2M phase genes        
g2m_genes <- cell_cycle_markers %>%
        dplyr::filter(phase == "G2/M") %>%
        pull("gene_name")
```

### 6.2.3 Cell Cycle Scoring

```{r}
# Perform cell cycle scoring
seurat_d60_bal_filtered_normalized <- CellCycleScoring(seurat_d60_bal_filtered_normalized,
 g2m.features = g2m_genes,
 s.features = s_genes)


```

### 6.2.4 Violin plot of effect cell cycle

```{r}
#| fig-width: 10
#| fig-height: 8
VlnPlot(seurat_d60_bal_filtered_normalized, features = c("S.Score", "G2M.Score"), group.by = "sampletag_name")
```

The violin plot shows clear evidence of a substantial amount of cycling genes in this case. Further investigate this with a PCA plot

### 6.2.5 PCA plot to assess effect cell cycle

To do PCA we have to do 2 things, first choose the most variable genes (features) and then scale the data. This order is important because, highly expressed genes exhibit the highest amount of variation and we don't want our 'highly variable genes' only to reflect high expression, we need to scale the data to scale variation with expression level.

```{r}
seurat_d60_bal_filtered_normalized <- FindVariableFeatures(seurat_d60_bal_filtered_normalized, 
                     selection.method = "vst",
                     nfeatures = 2000, 
                     verbose = FALSE)
seurat_d60_bal_filtered_normalized = ScaleData(seurat_d60_bal_filtered_normalized)
seurat_d60_bal_filtered_normalized = RunPCA(seurat_d60_bal_filtered_normalized)

```

```{r}
#| fig-width: 10
#| fig-height: 8
DimPlot(seurat_d60_bal_filtered_normalized,
        reduction = "pca",
        group.by= "Phase",
        split.by = "Phase")
```

```{r}
#| fig-width: 10
#| fig-height: 8
DimPlot(seurat_d60_bal_filtered_normalized,
        reduction = "pca",
        group.by= "Phase")
```

# Clear effect of cell cycle

```{r}
seurat_d60_bal_filtered_normalized |> write_rds("intermediate_data/seurat_obj_experiment_1_bal_afterQCdashboard.rds")
```

## 6.3 Check for effect of sex

```{r}
y_chromosome_genes<-   c("Ddx3y", "Eif2s3y", "Kdm5d")
seurat_d60_bal_filtered_normalized$gender <- seurat_d60_bal_filtered_normalized |> 
  join_features(features=c(y_chromosome_genes, "Xist"), slot="data", shape="wide") |>
  #as_tibble() |> 
  rowwise() |> 
  mutate(max_male_genes=max(c_across(Ddx3y:Kdm5d))
  ) |> 
  mutate(gender=case_when(
    (Xist==0)&(max_male_genes==0)~"no_expression_of_specific_genes",
    (Xist>0)&(max_male_genes==0)~"female",
    (Xist==0)&(max_male_genes>0)~"male",
    (Xist>0)&(max_male_genes>0)~"detection_of_male_and_female_genes",
    TRUE~"something is wrong")) |> 
  pull(gender)
```

```{r}
#| fig-width: 20
#| fig-height: 16
DimPlot(seurat_d60_bal_filtered_normalized,
        reduction = "pca",
        group.by= "gender", split.by = "gender")
```

```{r}
#| fig-width: 10
#| fig-height: 8
DimPlot(seurat_d60_bal_filtered_normalized,
        reduction = "pca",
        group.by= "gender")
```
