---
title: "Questions_ Prof_Machiels_18/03/2024"
format: 
  html:
    code-fold: true
editor: visual
toc: true
execute:
  warning: false
  message: false
---

```{r}
easypackages::libraries("viridis","patchwork", "Seurat", "tidyverse","tidyseurat", "DESeq2", "ggrepel")
obj.v5 <- read_rds("../../Documents/machiels_lab_viral/intermediate_data/seurat_obj_central.rds")
```

## select only clusters 0 1 6 7

```{r}
alveolar_clusters = c(0,1,6,7)
seurat_alveolar = subset(obj.v5, idents = c(0,1,6,7))
```

## remove day 8 since we only want to compare between mock PVM and MuHV4 in MS4A3+ and MS4A3- in lung and bal

```{r}
seurat_nod8 = obj.v5 %>%filter(!str_detect(day_mock, "d8"))
#d60
seurat_d60 <- seurat_nod8 %>%filter(str_detect(day_mock, "d60"))
seurat_onlyMuHV4_d60 <- seurat_d60 %>% filter(str_detect(condition, "MuHV4"))
seurat_onlyMuHV4_d60_bal <- seurat_onlyMuHV4_d60 %>%  filter(str_detect(sample_type, "bal"))

#mock
seurat_mock <- seurat_nod8 %>%filter(str_detect(day_mock, "Mock"))
seurat_mock_onlyMuHV4 = seurat_mock %>%  filter(str_detect(condition, "MuHV4"))
seurat_mock_onlyMuHV4_bal = seurat_mock_onlyMuHV4 %>% filter(str_detect(sample_type, "bal"))
```

## create seurat objects specific for Ms4a3 condition

```{r}
seurat_alveolar_ms4a3pos = seurat_alveolar %>%filter(!str_detect(sampletag_Ms4a3, "Ms4a3_neg"))
seurat_alveolar_ms4a3pos_noPR8 = seurat_alveolar %>%filter(!str_detect(virus, "PR8"))
seurat_alveolar_ms4a3pos = seurat_alveolar_ms4a3pos_noPR8 %>%filter(!str_detect(virus, "MAV1"))


seurat_alveolar_ms4a3neg = seurat_alveolar %>%filter(!str_detect(sampletag_Ms4a3, "Ms4a3_pos"))
seurat_alveolar_ms4a3neg_noPR8 = seurat_alveolar %>%filter(!str_detect(virus, "PR8"))
seurat_alveolar_ms4a3neg = seurat_alveolar_ms4a3neg_noPR8 %>%filter(!str_detect(virus, "MAV1"))
```

## create plots for MS4A3 +

```{r}
featureplot_alveolar_pparg_pos = FeaturePlot(seurat_alveolar_ms4a3pos, features = c("Pparg"), reduction = "umap.harmony_8dims", label = T, label.size = , split.by = c("virus"), cols = c("green", "red")) & theme(legend.position = "right")
featureplot_alveolar_pparg_pos
ggsave(filename = "featureplot_alveolar_pparg.png", width = 45, height = 20, units = "cm", path = "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/ms4a3pos/")
```

```{r}
featureplot_alveolar_cd11c_pos = FeaturePlot(seurat_alveolar_ms4a3pos, features = c("Cd11c"), reduction = "umap.harmony_8dims", label = T, label.size = , split.by = c("virus"), cols = c("green", "red")) & theme(legend.position = "right")
featureplot_alveolar_pparg_pos
ggsave(filename = "featureplot_alveolar_cd11c.png", width = 45, height = 20, units = "cm", path = "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/ms4a3pos/")
```

```{r}
featureplot_alveolar_h2_pos = FeaturePlot(seurat_alveolar_ms4a3pos, features = c("H2-ia-ie-AbSeq"), reduction = "umap.harmony_8dims", label = T, label.size = , split.by = c("virus"), cols = c("green", "red")) & theme(legend.position = "right")
featureplot_alveolar_h2_pos
ggsave(filename = "featureplot_alveolar_h2.png", width = 45, height = 20, units = "cm", path = "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/ms4a3pos/")
```

```{r}
featureplot_alveolar_siglec_pos = FeaturePlot(seurat_alveolar_ms4a3pos, features = c("Siglecf-AbSeq"), reduction = "umap.harmony_8dims", label = T, label.size = , split.by = c("virus"), cols = c("green", "red")) & theme(legend.position = "right")
featureplot_alveolar_siglec_pos
ggsave(filename = "featureplot_alveolar_siglec.png", width = 45, height = 20, units = "cm", path = "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/ms4a3pos/")
```

```{r}
featureplot_alveolar_cd274_pos = FeaturePlot(seurat_alveolar_ms4a3pos, features = c("Cd274-AbSeq"), reduction = "umap.harmony_8dims", label = T, label.size = , split.by = c("virus"), cols = c("green", "red")) & theme(legend.position = "right")
featureplot_alveolar_cd274_pos
ggsave(filename = "featureplot_alveolar_cd274.png", width = 45, height = 20, units = "cm", path = "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/ms4a3pos/")
```

## create plots for MS4A3 -

```{r}
featureplot_alveolar_pparg_neg = FeaturePlot(seurat_alveolar_ms4a3neg, features = c("Pparg"), reduction = "umap.harmony_8dims", label = T, label.size = , split.by = c("virus"), cols = c("green", "red")) & theme(legend.position = "right")
featureplot_alveolar_pparg_neg
ggsave(filename = "featureplot_alveolar_pparg.png", width = 45, height = 20, units = "cm", path = "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/ms4a3neg/")
```

```{r}
featureplot_alveolar_cd11c_neg = FeaturePlot(seurat_alveolar_ms4a3neg, features = c("Cd11c"), reduction = "umap.harmony_8dims", label = T, label.size = , split.by = c("virus"), cols = c("green", "red")) & theme(legend.position = "right")
featureplot_alveolar_pparg_neg
ggsave(filename = "featureplot_alveolar_cd11c.png", width = 45, height = 20, units = "cm", path = "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/ms4a3neg/")
```

```{r}
featureplot_alveolar_h2_neg = FeaturePlot(seurat_alveolar_ms4a3neg, features = c("H2-ia-ie-AbSeq"), reduction = "umap.harmony_8dims", label = T, label.size = , split.by = c("virus"), cols = c("green", "red")) & theme(legend.position = "right")
featureplot_alveolar_h2_neg
ggsave(filename = "featureplot_alveolar_h2.png", width = 45, height = 20, units = "cm", path = "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/ms4a3neg/")
```

```{r}
featureplot_alveolar_siglec_neg = FeaturePlot(seurat_alveolar_ms4a3neg, features = c("Siglecf-AbSeq"), reduction = "umap.harmony_8dims", label = T, label.size = , split.by = c("virus"), cols = c("green", "red")) & theme(legend.position = "right")
featureplot_alveolar_siglec_neg
ggsave(filename = "featureplot_alveolar_siglec.png", width = 45, height = 20, units = "cm", path = "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/ms4a3neg/")
```

```{r}
featureplot_alveolar_cd274_neg = FeaturePlot(seurat_alveolar_ms4a3neg, features = c("Cd274-AbSeq"), reduction = "umap.harmony_8dims", label = T, label.size = , split.by = c("virus"), cols = c("green", "red")) & theme(legend.position = "right")
featureplot_alveolar_cd274_neg
ggsave(filename = "featureplot_alveolar_cd274.png", width = 45, height = 20, units = "cm", path = "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/ms4a3neg/")
```

## Create plots for Ms4a3 + bal

```{r}
seurat_alveolar_ms4a3pos_bal <- seurat_alveolar_ms4a3pos %>%filter(str_detect(sample_type, "bal"))

```

```{r}
featureplot_alveolar_pparg_pos_bal = FeaturePlot(seurat_alveolar_ms4a3pos_bal, features = c("Pparg"), reduction = "umap.harmony_8dims", label = T, label.size = , split.by = c("virus"), cols = c("green", "red")) & theme(legend.position = "right")
featureplot_alveolar_pparg_pos_bal
ggsave(filename = "featureplot_alveolar_pparg.png", width = 45, height = 20, units = "cm", path = "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/ms4a3pos/bal/")
```

```{r}
featureplot_alveolar_cd11c_pos_bal = FeaturePlot(seurat_alveolar_ms4a3pos_bal, features = c("Cd11c"), reduction = "umap.harmony_8dims", label = T, label.size = , split.by = c("virus"), cols = c("green", "red")) & theme(legend.position = "right")
featureplot_alveolar_pparg_pos_bal
ggsave(filename = "featureplot_alveolar_cd11c.png", width = 45, height = 20, units = "cm", path = "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/ms4a3pos/bal/")
```

```{r}
featureplot_alveolar_h2_pos_bal = FeaturePlot(seurat_alveolar_ms4a3pos_bal, features = c("H2-ia-ie-AbSeq"), reduction = "umap.harmony_8dims", label = T, label.size = , split.by = c("virus"), cols = c("green", "red")) & theme(legend.position = "right")
featureplot_alveolar_h2_pos_bal
ggsave(filename = "featureplot_alveolar_h2.png", width = 45, height = 20, units = "cm", path = "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/ms4a3pos/bal/")
```

```{r}
featureplot_alveolar_siglec_pos_bal = FeaturePlot(seurat_alveolar_ms4a3pos_bal, features = c("Siglecf-AbSeq"), reduction = "umap.harmony_8dims", label = T, label.size = , split.by = c("virus"), cols = c("green", "red")) & theme(legend.position = "right")
featureplot_alveolar_siglec_pos_bal
ggsave(filename = "featureplot_alveolar_siglec.png", width = 45, height = 20, units = "cm", path = "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/ms4a3pos/bal/")
```

```{r}
featureplot_alveolar_cd274_pos_bal = FeaturePlot(seurat_alveolar_ms4a3pos_bal, features = c("Cd274-AbSeq"), reduction = "umap.harmony_8dims", label = T, label.size = , split.by = c("virus"), cols = c("green", "red")) & theme(legend.position = "right")
featureplot_alveolar_cd274_pos_bal
ggsave(filename = "featureplot_alveolar_cd274.png", width = 45, height = 20, units = "cm", path = "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/ms4a3pos/bal/")
```

## Create plots for Ms4a3 + lung

```{r}
seurat_alveolar_ms4a3pos_lung <- seurat_alveolar_ms4a3pos %>%filter(str_detect(sample_type, "lung"))
```

```{r}
featureplot_alveolar_pparg_pos_lung = FeaturePlot(seurat_alveolar_ms4a3pos_lung, features = c("Pparg"), reduction = "umap.harmony_8dims", label = T, label.size = , split.by = c("virus"), cols = c("green", "red")) & theme(legend.position = "right")
featureplot_alveolar_pparg_pos_lung
ggsave(filename = "featureplot_alveolar_pparg.png", width = 45, height = 20, units = "cm", path = "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/ms4a3pos/lung/")
```

```{r}
featureplot_alveolar_cd11c_pos_lung = FeaturePlot(seurat_alveolar_ms4a3pos_lung, features = c("Cd11c"), reduction = "umap.harmony_8dims", label = T, label.size = , split.by = c("virus"), cols = c("green", "red")) & theme(legend.position = "right")
featureplot_alveolar_pparg_pos_lung
ggsave(filename = "featureplot_alveolar_cd11c.png", width = 45, height = 20, units = "cm", path = "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/ms4a3pos/lung/")
```

```{r}
featureplot_alveolar_h2_pos_lung = FeaturePlot(seurat_alveolar_ms4a3pos_lung, features = c("H2-ia-ie-AbSeq"), reduction = "umap.harmony_8dims", label = T, label.size = , split.by = c("virus"), cols = c("green", "red")) & theme(legend.position = "right")
featureplot_alveolar_h2_pos_lung
ggsave(filename = "featureplot_alveolar_h2.png", width = 45, height = 20, units = "cm", path = "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/ms4a3pos/lung/")
```

```{r}
featureplot_alveolar_siglec_pos_lung = FeaturePlot(seurat_alveolar_ms4a3pos_lung, features = c("Siglecf-AbSeq"), reduction = "umap.harmony_8dims", label = T, label.size = , split.by = c("virus"), cols = c("green", "red")) & theme(legend.position = "right")
featureplot_alveolar_siglec_pos_lung
ggsave(filename = "featureplot_alveolar_siglec.png", width = 45, height = 20, units = "cm", path = "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/ms4a3pos/lung/")
```

```{r}
featureplot_alveolar_cd274_pos_lung = FeaturePlot(seurat_alveolar_ms4a3pos_lung, features = c("Cd274-AbSeq"), reduction = "umap.harmony_8dims", label = T, label.size = , split.by = c("virus"), cols = c("green", "red")) & theme(legend.position = "right")
featureplot_alveolar_cd274_pos_lung
ggsave(filename = "featureplot_alveolar_cd274.png", width = 45, height = 20, units = "cm", path = "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/ms4a3pos/lung/")
```

## Create plots for Ms4a3 - bal

```{r}
seurat_alveolar_ms4a3neg_bal <- seurat_alveolar_ms4a3neg %>%filter(str_detect(sample_type, "bal"))
```

```{r}
featureplot_alveolar_pparg_neg_bal = FeaturePlot(seurat_alveolar_ms4a3neg_bal, features = c("Pparg"), reduction = "umap.harmony_8dims", label = T, label.size = , split.by = c("virus"), cols = c("green", "red")) & theme(legend.position = "right")
featureplot_alveolar_pparg_pos_bal
ggsave(filename = "featureplot_alveolar_pparg.png", width = 45, height = 20, units = "cm", path = "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/ms4a3neg/bal/")
```

```{r}
featureplot_alveolar_cd11c_neg_bal = FeaturePlot(seurat_alveolar_ms4a3neg_bal, features = c("Cd11c"), reduction = "umap.harmony_8dims", label = T, label.size = , split.by = c("virus"), cols = c("green", "red")) & theme(legend.position = "right")
featureplot_alveolar_pparg_pos_bal
ggsave(filename = "featureplot_alveolar_cd11c.png", width = 45, height = 20, units = "cm", path = "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/ms4a3neg/bal/")
```

```{r}
featureplot_alveolar_h2_neg_bal = FeaturePlot(seurat_alveolar_ms4a3neg_bal, features = c("H2-ia-ie-AbSeq"), reduction = "umap.harmony_8dims", label = T, label.size = , split.by = c("virus"), cols = c("green", "red")) & theme(legend.position = "right")
featureplot_alveolar_h2_pos_bal
ggsave(filename = "featureplot_alveolar_h2.png", width = 45, height = 20, units = "cm", path = "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/ms4a3neg/bal/")
```

```{r}
featureplot_alveolar_siglec_neg_bal = FeaturePlot(seurat_alveolar_ms4a3neg_bal, features = c("Siglecf-AbSeq"), reduction = "umap.harmony_8dims", label = T, label.size = , split.by = c("virus"), cols = c("green", "red")) & theme(legend.position = "right")
featureplot_alveolar_siglec_pos_bal
ggsave(filename = "featureplot_alveolar_siglec.png", width = 45, height = 20, units = "cm", path = "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/ms4a3neg/bal/")
```

```{r}
featureplot_alveolar_cd274_neg_bal = FeaturePlot(seurat_alveolar_ms4a3neg_bal, features = c("Cd274-AbSeq"), reduction = "umap.harmony_8dims", label = T, label.size = , split.by = c("virus"), cols = c("green", "red")) & theme(legend.position = "right")
featureplot_alveolar_cd274_pos_bal
ggsave(filename = "featureplot_alveolar_cd274.png", width = 45, height = 20, units = "cm", path = "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/ms4a3neg/bal/")
```

## Create plots for Ms4a3 - lung

```{r}
seurat_alveolar_ms4a3neg_lung <- seurat_alveolar_ms4a3neg %>%filter(str_detect(sample_type, "lung"))
```

```{r}
featureplot_alveolar_pparg_neg_lung = FeaturePlot(seurat_alveolar_ms4a3neg_lung, features = c("Pparg"), reduction = "umap.harmony_8dims", label = T, label.size = , split.by = c("virus"), cols = c("green", "red")) & theme(legend.position = "right")
featureplot_alveolar_pparg_pos_lung
ggsave(filename = "featureplot_alveolar_pparg.png", width = 45, height = 20, units = "cm", path = "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/ms4a3neg/lung/")
```

```{r}
featureplot_alveolar_cd11c_neg_lung = FeaturePlot(seurat_alveolar_ms4a3neg_lung, features = c("Cd11c"), reduction = "umap.harmony_8dims", label = T, label.size = , split.by = c("virus"), cols = c("green", "red")) & theme(legend.position = "right")
featureplot_alveolar_pparg_pos_lung
ggsave(filename = "featureplot_alveolar_cd11c.png", width = 45, height = 20, units = "cm", path = "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/ms4a3neg/lung/")
```

```{r}
featureplot_alveolar_h2_neg_lung = FeaturePlot(seurat_alveolar_ms4a3neg_lung, features = c("H2-ia-ie-AbSeq"), reduction = "umap.harmony_8dims", label = T, label.size = , split.by = c("virus"), cols = c("green", "red")) & theme(legend.position = "right")
featureplot_alveolar_h2_pos_lung
ggsave(filename = "featureplot_alveolar_h2.png", width = 45, height = 20, units = "cm", path = "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/ms4a3neg/lung/")
```

```{r}
featureplot_alveolar_siglec_neg_lung = FeaturePlot(seurat_alveolar_ms4a3neg_lung, features = c("Siglecf-AbSeq"), reduction = "umap.harmony_8dims", label = T, label.size = , split.by = c("virus"), cols = c("green", "red")) & theme(legend.position = "right")
featureplot_alveolar_siglec_pos_lung
ggsave(filename = "featureplot_alveolar_siglec.png", width = 45, height = 20, units = "cm", path = "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/ms4a3neg/lung/")
```

```{r}
featureplot_alveolar_cd274_neg_lung = FeaturePlot(seurat_alveolar_ms4a3neg_lung, features = c("Cd274-AbSeq"), reduction = "umap.harmony_8dims", label = T, label.size = , split.by = c("virus"), cols = c("green", "red")) & theme(legend.position = "right")
featureplot_alveolar_cd274_pos_lung
ggsave(filename = "featureplot_alveolar_cd274.png", width = 45, height = 20, units = "cm", path = "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/ms4a3neg/lung/")

```

## CD11c bal d60

```{r}
featureplot_d60_cd11c = FeaturePlot(seurat_onlyMuHV4_d60_bal, reduction = "umap.harmony_8dims", label = T, split.by = c("sampletag_Ms4a3"), features = "Cd11c", repel = TRUE)


featureplot_d60_cd11c 

# + ggtitle("",subtitle = "UMAP plots with expression of CD11c at day 60 between Ms4a3 conditions in  mouse bal") + theme(plot.title = element_text(hjust = 0.5))



ggsave(filename = "featureplot_d60_cd11c.png", width = 45, height = 20, units = "cm", path = "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/d60/")
```

```{r}
featureplot_mock_cd11c = FeaturePlot(seurat_mock_onlyMuHV4_bal, reduction = "umap.harmony_8dims", label = T,label.size = 5, split.by = c("sampletag_Ms4a3"), features = "Cd11c")
featureplot_mock_cd11c

# + ggtitle("",subtitle = "UMAP plots with expression of CD11c at day 0 (mock) between Ms4a3 conditions in  mouse bal") + theme(plot.title = element_text(hjust = 0.5))

ggsave(filename = "featureplot_mock_cd11c.png", width = 45, height = 20, units = "cm", path = "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/mock/")
```

## PPARy bal d60

```{r}
featureplot_d60_ppary = FeaturePlot(seurat_onlyMuHV4_d60_bal, reduction = "umap.harmony_8dims", label = T, split.by = c("sampletag_Ms4a3"), features = "Pparg", repel = TRUE)


featureplot_d60_ppary 

# + ggtitle("",subtitle = "UMAP plots with expression of CD11c at day 60 between Ms4a3 conditions in  mouse bal") + theme(plot.title = element_text(hjust = 0.5))



ggsave(filename = "featureplot_d60_ppary.png", width = 45, height = 20, units = "cm", path = "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/d60/")
```

```{r}
featureplot_mock_pparg = FeaturePlot(seurat_mock_onlyMuHV4_bal, reduction = "umap.harmony_8dims", label = T,label.size = 5, split.by = c("sampletag_Ms4a3"), features = "Pparg")
featureplot_mock_pparg

# + ggtitle("",subtitle = "UMAP plots with expression of CD11c at day 0 (mock) between Ms4a3 conditions in  mouse bal") + theme(plot.title = element_text(hjust = 0.5))

ggsave(filename = "featureplot_mock_pparg.png", width = 45, height = 20, units = "cm", path = "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/mock/")
```

## SiglecF bal d60

```{r}
featureplot_d60_siglecf = FeaturePlot(seurat_onlyMuHV4_d60_bal, reduction = "umap.harmony_8dims", label = T, split.by = c("sampletag_Ms4a3"), features = "Siglecf-AbSeq", repel = TRUE)


featureplot_d60_siglecf

# + ggtitle("",subtitle = "UMAP plots with expression of CD11c at day 60 between Ms4a3 conditions in  mouse bal") + theme(plot.title = element_text(hjust = 0.5))



ggsave(filename = "featureplot_d60_siglecf.png", width = 45, height = 20, units = "cm", path = "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/d60/")
```

```{r}
featureplot_mock_siglecf = FeaturePlot(seurat_mock_onlyMuHV4_bal, reduction = "umap.harmony_8dims", label = T,label.size = 5, split.by = c("sampletag_Ms4a3"), features = "Siglecf-AbSeq")
featureplot_mock_siglecf

# + ggtitle("",subtitle = "UMAP plots with expression of CD11c at day 0 (mock) between Ms4a3 conditions in  mouse bal") + theme(plot.title = element_text(hjust = 0.5))

ggsave(filename = "featureplot_mock_siglecf.png", width = 45, height = 20, units = "cm", path = "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/mock/")
```

## H2-ia-ie bal d60

```{r}
featureplot_d60_h2 = FeaturePlot(seurat_onlyMuHV4_d60_bal, reduction = "umap.harmony_8dims", label = T, split.by = c("sampletag_Ms4a3"), features = "H2-ia-ie-AbSeq", repel = TRUE)


featureplot_d60_h2

# + ggtitle("",subtitle = "UMAP plots with expression of CD11c at day 60 between Ms4a3 conditions in  mouse bal") + theme(plot.title = element_text(hjust = 0.5))



ggsave(filename = "featureplot_d60_h2.png", width = 45, height = 20, units = "cm", path = "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/d60/")
```

```{r}
featureplot_mock_h2 = FeaturePlot(seurat_mock_onlyMuHV4_bal, reduction = "umap.harmony_8dims", label = T, split.by = c("sampletag_Ms4a3"), features = "H2-ia-ie-AbSeq", repel = TRUE)


featureplot_mock_h2

# + ggtitle("",subtitle = "UMAP plots with expression of CD11c at day 60 between Ms4a3 conditions in  mouse bal") + theme(plot.title = element_text(hjust = 0.5))



ggsave(filename = "featureplot_mock_h2.png", width = 45, height = 20, units = "cm", path = "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/mock/")
```

## CD274 d60

```{r}
featureplot_d60_cd274 = FeaturePlot(seurat_onlyMuHV4_d60_bal, reduction = "umap.harmony_8dims", label = T, split.by = c("sampletag_Ms4a3"), features = "Cd274-AbSeq", repel = TRUE)


featureplot_d60_cd274

# + ggtitle("",subtitle = "UMAP plots with expression of CD11c at day 60 between Ms4a3 conditions in  mouse bal") + theme(plot.title = element_text(hjust = 0.5))



ggsave(filename = "featureplot_d60_cd274.png", width = 45, height = 20, units = "cm", path = "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/d60/")
```

```{r}
featureplot_mock_cd274 = FeaturePlot(seurat_mock_onlyMuHV4_bal, reduction = "umap.harmony_8dims", label = T, split.by = c("sampletag_Ms4a3"), features = "Cd274-AbSeq", repel = TRUE)


featureplot_mock_cd274

# + ggtitle("",subtitle = "UMAP plots with expression of CD11c at day 60 between Ms4a3 conditions in  mouse bal") + theme(plot.title = element_text(hjust = 0.5))



ggsave(filename = "featureplot_mock_cd274.png", width = 45, height = 20, units = "cm", path = "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/mock/")
```

## identify idendity of cluster 7

```{r}
# identify markers distinguising cluster 7 from clusters 0 1 6
Idents(object = obj.v5) <- "harmony_cluster_8dims"
DefaultAssay(obj.v5) <- "RNA"
# identifies markers for cluster 7 distinguishing themselves from the clusters we are more certain off they are alv macro
# a gene is considered a marker if at least 20% of cells expressing the gene in either of the two groups of cells 
# a gene is considered a marker if there is at least a 20% difference of expression between that gene between the groups
# DESeq2 test is used to estimate the differentially expressed genes between cluster 7 and 0,1,6

markersCluster7 = FindMarkers(obj.v5, ident.1 = 7, ident.2 = c(0,1,6),
                              min.diff.pct = 0.20, test.use = "DESeq2", min.pct = 0.2)
write.csv(markersCluster7, "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/markers_cluster7.csv")
# filter results, find those genes with a big difference between pct1 and pc2 who also have a high log fold change
markersCluster7$pct_difference = markersCluster7$pct.1 - markersCluster7$pct.2
markersCluster7_filtered = markersCluster7[markersCluster7$pct_difference > 0.25,]
markersCluster7_filtered = markersCluster7_filtered[markersCluster7_filtered$avg_log2FC > 1,]
markersCluster7_filtered = cbind(rownames(markersCluster7_filtered), markersCluster7_filtered)

write.csv(markersCluster7_filtered, "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/markers_cluster7_filtered.csv")

markersCluster7_filtered_genes = markersCluster7_filtered$gene
markersCluster7_filtered_genes
write.table(markersCluster7_filtered_genes,  "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/markers_cluster7_genes.txt")
```

## UMAPs showing **RNA data** (not Abseq expression)  for the bal cells at day60 and mock, for **PVM** (not MuHV-4) in the clusters corresponding to the alveolar macrophages

```{r}
seurat_onlyPVM =seurat_nod8 %>% filter(str_detect(condition, "PVM"))
seurat_onlyPVM_bal =seurat_nod8 %>% filter(str_detect(sample_type, "bal"))
seurat_onlyPVM_bal_alveolar = subset(seurat_onlyPVM_bal, idents = c(0,1,6,7))
DefaultAssay(seurat_onlyPVM_bal_alveolar) <- "RNA"

```

```{r}
featureplot_pvm_alveolar_pparg =  FeaturePlot(seurat_onlyPVM_bal_alveolar, reduction = "umap.harmony_8dims", label = T, split.by = c("day_mock"), features = "Pparg")
featureplot_pvm_alveolar_pparg
ggsave(filename = "featureplot_alveolar_pvm_ppag.png", width = 45, height = 20, units = "cm", path = "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/alveolar/")
```

```{r}
featureplot_pvm_alveolar_siglec =  FeaturePlot(seurat_onlyPVM_bal_alveolar, reduction = "umap.harmony_8dims", label = T, split.by = c("day_mock"), features = "Siglecf")
featureplot_pvm_alveolar_siglec
ggsave(filename = "featureplot_alveolar_pvm_siglecf.png", width = 45, height = 20, units = "cm", path = "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/alveolar/")
```

```{r}
featureplot_pvm_alveolar_cd274 =  FeaturePlot(seurat_onlyPVM_bal_alveolar, reduction = "umap.harmony_8dims", label = T, split.by = c("day_mock"), features = "Cd274")
featureplot_pvm_alveolar_cd274
ggsave(filename = "featureplot_alveolar_pvm_cd274.png", width = 45, height = 20, units = "cm", path = "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/alveolar/")
```

```{r}
featureplot_pvm_alveolar_cd11c =  FeaturePlot(seurat_onlyPVM_bal_alveolar, reduction = "umap.harmony_8dims", label = T, split.by = c("day_mock"), features = "Itgax")
featureplot_pvm_alveolar_cd11c
ggsave(filename = "featureplot_alveolar_pvm_cd11c.png", width = 45, height = 20, units = "cm", path = "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/alveolar/")
```

## Identify DE genes between MS4A3 + and MS4A3 - in cluster1 of d60 BAL PVM mice

significantly differentially expressed genes

```{r}
seurat_cluster1 = subset(seurat_onlyPVM_bal_alveolar, idents = 1)
DefaultAssay(seurat_cluster1) <- "RNA"
Idents(seurat_cluster1) <- "sampletag_Ms4a3"
# a gene is considered a marker if at least 20% of cells expressing the gene in either of the two groups of cells 
# a gene is considered a marker if there is at least a 20% difference of expression between that gene between the groups
# DESeq2 test is used to estimate the differentially expressed genes between cluster 7 and 0,1,6
markersCluster1_Ms4a3neg = FindMarkers(seurat_cluster1, ident.1 = "Ms4a3_neg" , 
                                       ident.2 = "Ms4a3_pos",
                                       min.diff.pct = 0.20, test.use = "DESeq2", 
                                       min.pct = 0.2)
markersCluster1_Ms4a3pos = FindMarkers(seurat_cluster1, ident.1 = "Ms4a3_pos" , 
                                       ident.2 = "Ms4a3_neg",
                                       min.diff.pct = 0.20, test.use = "DESeq2", 
                                       min.pct = 0.2)
```

### the differentially expressed genes have an average log2 fold change over 1 and are in 10% of the cells of Ms4a3 negative group more expressed compared to the Ms4a3 positive group

```{r}
markersCluster1_Ms4a3neg = cbind(rownames(markersCluster1_Ms4a3neg), markersCluster1_Ms4a3neg)
markersCluster1_Ms4a3neg$pct_difference = markersCluster1_Ms4a3neg$pct.1 - markersCluster1_Ms4a3neg$pct.2
markersCluster1_Ms4a3neg_filtered = markersCluster1_Ms4a3neg[markersCluster1_Ms4a3neg$pct_difference > 0.1,]
markersCluster1_Ms4a3neg_filtered = markersCluster1_Ms4a3neg_filtered[markersCluster1_Ms4a3neg_filtered$avg_log2FC > 1,]
markersCluster1_Ms4a3neg_filtered = cbind(rownames(markersCluster1_Ms4a3neg_filtered), markersCluster1_Ms4a3neg_filtered)

write.csv(markersCluster1_Ms4a3neg, "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/markers_Cluster1_Ms4a3neg.csv")

write.csv(markersCluster1_Ms4a3neg_filtered, "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/markers_Cluster1_Ms4a3neg_filtered.csv")

markersCluster1_Ms4a3neg_filtered_genes = markersCluster1_Ms4a3neg_filtered$`rownames(markersCluster1_Ms4a3neg_filtered)`
markersCluster1_Ms4a3neg_filtered_genes
write.table(markersCluster1_Ms4a3neg_filtered_genes,  "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/markers_Cluster1_Ms4a3neg_genes.txt")
```

### the differentially expressed genes have an average log2 fold change over 1 and are in 10% of the cells of Ms4a3 positive group more expressed compared to the Ms4a3 negative group

```{r}
markersCluster1_Ms4a3pos = cbind(rownames(markersCluster1_Ms4a3pos), markersCluster1_Ms4a3pos)
markersCluster1_Ms4a3pos$pct_difference = markersCluster1_Ms4a3pos$pct.1 - markersCluster1_Ms4a3pos$pct.2
markersCluster1_Ms4a3pos_filtered = markersCluster1_Ms4a3pos[markersCluster1_Ms4a3pos$pct_difference > 0.1,]
markersCluster1_Ms4a3pos_filtered = markersCluster1_Ms4a3pos_filtered[markersCluster1_Ms4a3pos_filtered$avg_log2FC > 1,]
markersCluster1_Ms4a3pos_filtered = cbind(rownames(markersCluster1_Ms4a3pos_filtered), markersCluster1_Ms4a3pos_filtered)

write.csv(markersCluster1_Ms4a3pos, "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/markers_Cluster1_Ms4a3pos.csv")
write.csv(markersCluster1_Ms4a3pos_filtered, "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/markers_Cluster1_Ms4a3pos_filtered.csv")

markersCluster1_Ms4a3pos_filtered_genes = markersCluster1_Ms4a3pos_filtered$`rownames(markersCluster1_Ms4a3pos_filtered)`
markersCluster1_Ms4a3pos_filtered_genes
write.table(markersCluster1_Ms4a3pos_filtered_genes,  "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/markers_Cluster1_Ms4a3pos_genes.txt")
```

```{r}
markersCluster1_Ms4a3neg$diffexpressed <- "NO"
markersCluster1_Ms4a3neg$diffexpressed[markersCluster1_Ms4a3neg$avg_log2FC > 1 & markersCluster1_Ms4a3neg$pct_difference > 0.1] <- "UP"

markersCluster1_Ms4a3neg$diffexpressed[markersCluster1_Ms4a3neg$avg_log2FC < -1 & markersCluster1_Ms4a3neg$pct_difference < -0.1] <- "DOWN"

markersCluster1_Ms4a3neg$delabel <- NA

markersCluster1_Ms4a3neg$delabel[markersCluster1_Ms4a3neg$diffexpressed != "NO"] <- markersCluster1_Ms4a3neg$`rownames(markersCluster1_Ms4a3neg)`[markersCluster1_Ms4a3neg$diffexpressed != "NO"]



volcano_Ms4a3neg <- ggplot(data=markersCluster1_Ms4a3neg,
                           aes(x=avg_log2FC, y=-log10(p_val_adj), col = diffexpressed,                            label = delabel)) +
                           geom_point() + theme_minimal() +
                           theme(panel.grid = element_blank()) +
                           scale_x_continuous(breaks = seq(floor(min(markersCluster1_Ms4a3neg$avg_log2FC)), 
                                            ceiling(max(markersCluster1_Ms4a3neg$avg_log2FC)), by = 0.5)) +
                           geom_text_repel(max.overlaps = 100)
volcano_Ms4a3neg
ggsave("volcano_Ms4a3neg.png",width = 45, height = 20, units = "cm", path = "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/volcano")
```

```{r}
markersCluster1_Ms4a3pos$diffexpressed <- "NO"
markersCluster1_Ms4a3pos$diffexpressed[markersCluster1_Ms4a3pos$avg_log2FC > 1 & markersCluster1_Ms4a3pos$pct_difference > 0.1] <- "UP"

markersCluster1_Ms4a3pos$diffexpressed[markersCluster1_Ms4a3pos$avg_log2FC < -1 & markersCluster1_Ms4a3pos$pct_difference < -0.1] <- "DOWN"

markersCluster1_Ms4a3pos$delabel <- NA

markersCluster1_Ms4a3pos$delabel[markersCluster1_Ms4a3pos$diffexpressed != "NO"] <- markersCluster1_Ms4a3pos$`rownames(markersCluster1_Ms4a3pos)`[markersCluster1_Ms4a3pos$diffexpressed != "NO"]



volcano_Ms4a3pos <- ggplot(data=markersCluster1_Ms4a3pos,
                           aes(x=avg_log2FC, y=-log10(p_val_adj), col = diffexpressed,                            label = delabel)) +
                           geom_point() + theme_minimal() +
                           theme(panel.grid = element_blank()) +
                           scale_x_continuous(breaks = seq(floor(min(markersCluster1_Ms4a3pos$avg_log2FC)), 
                                            ceiling(max(markersCluster1_Ms4a3pos$avg_log2FC)), by = 0.5)) +
                           geom_text_repel(max.overlaps = 100)
volcano_Ms4a3pos
ggsave("volcano_Ms4a3pos.png",width = 45, height = 20, units = "cm", path = "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/volcano")
```

## Identify DE genes between mock and PVM in cluster1 of d60 BAL PVM mice

### 

```{r}

seurat_nod8_PVMMock = seurat_nod8 %>% filter(!str_detect(condition, "MuHV4"))
seurat_nod8_PVMMock = seurat_nod8_PVMMock %>% filter(!str_detect(condition, "PR8"))
seurat_nod8_PVMMock = seurat_nod8_PVMMock %>% filter(!str_detect(condition, "MAV1"))
seurat_nod8_PVMMock_bal = seurat_nod8_PVMMock %>% filter(str_detect(sample_type, "bal"))
seurat_nod8_PVMMock_bal =  seurat_nod8_PVMMock_bal %>% filter(str_detect(day_mock, "d60"))
seurat_nod8_PVMMock_bal_cluster1 = subset(seurat_nod8_PVMMock_bal, idents = 1)
DefaultAssay(seurat_nod8_PVMMock_bal_cluster1) <- "RNA"

Idents(seurat_nod8_PVMMock_bal_cluster1) <- "condition"

# a gene is considered a marker if at least 20% of cells expressing the gene in either of the two groups of cells 
# a gene is considered a marker if there is at least a 20% difference of expression between that gene between the groups
# DESeq2 test is used to estimate the differentially expressed genes between cluster 7 and 0,1,6
markersCluster1_Mock = FindMarkers(seurat_nod8_PVMMock_bal_cluster1, ident.1 = "Mock",
                                       ident.2 = "PVM",
                                       min.diff.pct = 0.20, test.use = "DESeq2", 
                                       min.pct = 0.2)
markersCluster1_PVM = FindMarkers(seurat_nod8_PVMMock_bal_cluster1, ident.1 = "PVM" , 
                                       ident.2 = "Mock",
                                       min.diff.pct = 0.20, test.use = "DESeq2", 
                                       min.pct = 0.2)

```

### the differentially expressed genes have an average log2 fold change over 1 and are in 10% of the cells of Mock group more expressed compared to the PVM group

```{r}
markersCluster1_Mock = cbind(rownames(markersCluster1_Mock), markersCluster1_Mock)
markersCluster1_Mock$pct_difference = markersCluster1_Mock$pct.1 - markersCluster1_Mock$pct.2
markersCluster1_Mock_filtered = markersCluster1_Mock[markersCluster1_Mock$pct_difference > 0.1,]
markersCluster1_Mock_filtered = markersCluster1_Mock_filtered[markersCluster1_Mock_filtered$avg_log2FC > 1,]
markersCluster1_Mock_filtered = cbind(rownames(markersCluster1_Mock_filtered), markersCluster1_Mock_filtered)
write.csv(markersCluster1_Mock, "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/markers_Cluster1_Mock.csv")
write.csv(markersCluster1_Mock_filtered, "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/markers_Cluster1_Mock_filtered.csv")

markersCluster1_Mock_filtered_genes = markersCluster1_Mock_filtered$`rownames(markersCluster1_Mock_filtered)`
markersCluster1_Mock_filtered_genes
write.table(markersCluster1_Mock_filtered_genes,  "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/markers_Cluster1_Mock_genes.txt")
```

### the differentially expressed genes have an average log2 fold change over 1 and are in 10% of the cells of PVM group more expressed compared to the Mock group

```{r}
markersCluster1_PVM = cbind(rownames(markersCluster1_PVM), markersCluster1_PVM)
markersCluster1_PVM <- markersCluster1_PVM[, -c(2)]
markersCluster1_PVM$pct_difference = markersCluster1_PVM$pct.1 - markersCluster1_PVM$pct.2
markersCluster1_PVM_filtered = markersCluster1_PVM[markersCluster1_PVM$pct_difference > 0.1,]
markersCluster1_PVM_filtered = markersCluster1_PVM_filtered[markersCluster1_PVM_filtered$avg_log2FC > 1,]
markersCluster1_PVM_filtered = cbind(rownames(markersCluster1_PVM_filtered), markersCluster1_PVM_filtered)
write.csv(markersCluster1_PVM, "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/markers_Cluster1_PVM.csv")
write.csv(markersCluster1_PVM_filtered, "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/markers_Cluster1_PVM_filtered.csv")

markersCluster1_PVM_filtered_genes = markersCluster1_PVM_filtered$`rownames(markersCluster1_PVM_filtered)`
markersCluster1_PVM_filtered_genes
write.table(markersCluster1_PVM_filtered_genes,  "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/markers_Cluster1_PVM_genes.txt")
```

for visualization purposes are only those genes marked who are 20% or more expressed in the PVM or mock condition

```{r}
markersCluster1_Mock$diffexpressed <- "NO"
markersCluster1_Mock$diffexpressed[markersCluster1_Mock$avg_log2FC > 1 & markersCluster1_Mock$pct_difference > 0.2] <- "UP"

markersCluster1_Mock$diffexpressed[markersCluster1_Mock$avg_log2FC < -1 & markersCluster1_Mock$pct_difference < -0.2] <- "DOWN"

markersCluster1_Mock$delabel <- NA

markersCluster1_Mock$delabel[markersCluster1_Mock$diffexpressed != "NO"] <- markersCluster1_Mock$`rownames(markersCluster1_Mock)`[markersCluster1_Mock$diffexpressed != "NO"]



volcano_mock <- ggplot(data=markersCluster1_Mock,
                           aes(x=avg_log2FC, y=-log10(p_val_adj), col = diffexpressed,                            label = delabel)) +
                           geom_point() + theme_minimal() +
                           theme(panel.grid = element_blank()) +
                           scale_x_continuous(breaks = seq(floor(min(markersCluster1_Mock$avg_log2FC)), 
                                            ceiling(max(markersCluster1_Mock$avg_log2FC)), by = 0.5)) +
                           geom_text_repel(max.overlaps = 100)
volcano_mock
ggsave("volcano_mock.png",width = 45, height = 20, units = "cm", path = "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/volcano")
```

```{r}
markersCluster1_PVM$diffexpressed <- "NO"
markersCluster1_PVM$diffexpressed[markersCluster1_PVM$avg_log2FC > 1 & markersCluster1_PVM$pct_difference > 0.2] <- "UP"

markersCluster1_PVM$diffexpressed[markersCluster1_PVM$avg_log2FC < -1 & markersCluster1_PVM$pct_difference < -0.2] <- "DOWN"

markersCluster1_PVM$delabel <- NA

markersCluster1_PVM$delabel[markersCluster1_PVM$diffexpressed != "NO"] <- markersCluster1_PVM$`rownames(markersCluster1_PVM)`[markersCluster1_PVM$diffexpressed != "NO"]



volcano_PVM <- ggplot(data=markersCluster1_PVM,
                           aes(x=avg_log2FC, y=-log10(p_val_adj), col = diffexpressed,                            label = delabel)) +
                           geom_point() + theme_minimal() +
                           theme(panel.grid = element_blank()) +
                           scale_x_continuous(breaks = seq(floor(min(markersCluster1_PVM$avg_log2FC)), 
                                            ceiling(max(markersCluster1_PVM$avg_log2FC)), by = 0.5)) +
                           geom_text_repel(max.overlaps = 100)
volcano_PVM
ggsave("volcano_PVM.png",width = 45, height = 20, units = "cm", path = "../../Documents/machiels_lab_viral/output/plotsUmap/questionsBene19March/volcano")
```
